---
title: "Development actions history"
output: html_document
editor_options: 
  chunk_output_type: console
---

All commands that you use to use when developing packages...

# First time just after creating the project

- Fill the following chunk to create the DESCRIPTION of your package

```{r description}
# Describe your package
fusen::fill_description(
  pkg = here::here(),
  fields = list(
    Title = "Recent trends of nursery quality",
    Description = "Recent trends of nursery quality related parameters in French estuaries.",
    `Authors@R` = c(
      person("Florence", "Mounier", email = "florence.mounier@inrae.fr", role = c("aut", "cre"), comment = c(ORCID = "0000-0001-8916-7007"))
    )
  )
)
# Define License with use_*_license()
usethis::use_mit_license("Florence Mounier")
```

# All-in-one function to deploy publicly on GitHub

Either use this GitHub all-in-one function or choose the steps one-by-one in the following sections.
See `vignette("share-on-a-github-website", package = "fusen")` for more information.

```{r, eval=FALSE}
# _All-in-one share of your package and its documentation on GitHub
fusen::init_share_on_github()
# The next time you run this function, you can set `init_share_on_github(ask = TRUE)`, to not see all the intermediate questions.
```


# Start using git

```{r, eval=FALSE}
usethis::use_git()
# Deal with classical files to ignore
usethis::git_vaccinate()
# Use main for primary branch
usethis::git_default_branch_rename()
```

# Set extra sources of documentation

```{r, eval=FALSE}
# Install a first time
remotes::install_local()
# README
usethis::use_readme_rmd()
# Code of Conduct
# usethis::use_code_of_conduct("contact@fake.com")
# NEWS
usethis::use_news_md()
```

**From now, you will need to "inflate" your package at least once to be able to use the following commands. Let's go to your flat template, and come back here later if/when needed.**


# Package development tools
## Use once

```{r, eval=FALSE}
# Pipe
usethis::use_pipe()

# package-level documentation
usethis::use_package_doc()

# GitHub
# Add your credentials for GitHub
usethis::create_github_token()
gitcreds::gitcreds_set()
```

## Use everytime needed

```{r, eval=FALSE}
# Simulate package installation
pkgload::load_all()

# Generate documentation and deal with dependencies
attachment::att_amend_desc()

# Check the package
devtools::check()

# Global variables
checkhelper::print_globals()

## Add internal datasets ----
usethis::use_data_raw()
usethis::use_data()
checkhelper::use_data_doc()

## Vignette ----
usethis::use_vignette()
devtools::build_vignettes()

# Add a new flat template
fusen::add_flat_template("add")

# Inflate all flat files at once
fusen::inflate_all()

# Package structure (add this in a dev/Readme.md file)
fusen::draw_package_structure()
```

# Share the package

```{r, eval=FALSE}
# set and try pkgdown documentation website locally
usethis::use_pkgdown()
pkgdown::build_site()

# build the tar.gz with vignettes to share with others
devtools::build(vignettes = TRUE)

# Share your package and its documentation on GitHub
# usethis::create_github_token()
# gitcreds::gitcreds_set()
fusen::init_share_on_github()
```

