---
title: "data_contamination_metals"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{data_contamination_metals}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r libraries, include = FALSE}
library(esteem.overview)
library(tidyverse)
library(here)
```

# Get raw data
```{r}
data <- esteem.overview::data_ROCCHMV_cleaned |> ungroup()
```

# Data for metals
```{r}
metals <- data.frame(
  PARAMETRE_LIBELLE = c("Mercury", "Cadmium", "Lead", "Copper"),
  EC_MPC_µg_gww = c(0.5, 1, 1.5, NA)
  )

data_metals <- data |> 
  filter(support == "Bivalve") |> 
  select(-c(SUPPORT_NIVEAU_PRELEVEMENT, support, niveau)) |> 
  mutate(PARAMETRE_LIBELLE = case_when(
    PARAMETRE_LIBELLE == "Mercure" ~ "Mercury",
    PARAMETRE_LIBELLE == "Plomb" ~ "Lead",
    PARAMETRE_LIBELLE == "Cuivre" ~ "Copper",
    TRUE ~ PARAMETRE_LIBELLE
  )) |> 
  filter(PARAMETRE_LIBELLE %in% metals$PARAMETRE_LIBELLE) 
```

## Importance of the species matrix

Only for Bivalves (no interesting results for fish)

- Gironde : only C. gigas
- Seine : only M. edulis
- Loire : 
  - 1979-2021: M. edulis
  - 2022-2024: M. edulis + M.galloprovincialis
  - 2011-2013 : R. philippinarum duplicate
  - 2017-2024 : C. gigas
```{r}
data_metals |> 
  distinct(ESTUARY, espece)

data_metals |> 
  filter(ESTUARY == "Loire") |> 
  count(YEAR, espece)
```

For copper oyster and mussels bioaccumulation is very different
```{r}
data_metals |> 
  filter(ESTUARY == "Loire") |>
  ggplot() +
  aes(x = YEAR, y = RESULTAT, colour = espece) +
  geom_line() +
  facet_wrap(vars(PARAMETRE_LIBELLE), scales = "free_y")
```

Estimate species conversion factor from data in the Loire estuary (2017-2024)
```{r}
data_metals <- data_metals |> 
  filter(espece != "Ruditapes philippinarum (palourde japonaise)") |>
  mutate(SPECIES = case_when(
    espece == "Crassostrea gigas (huître creuse)" ~ "oyster",
    espece %in% c("Mytilus edulis (moule commune)",
                  "Mytilus edulis + galloprovincialis (moule)") ~ "mussel"
  ))

data_metals_loire <- data_metals |> 
  filter(ESTUARY == "Loire", YEAR >= 2017)
  
factors_metals <- data_metals_loire |> 
  group_by(PARAMETRE_LIBELLE, SPECIES) |> 
  summarise(median_RESULTAT = median(RESULTAT, na.rm = TRUE), .groups = "drop")  |> 
  pivot_wider(names_from = SPECIES, values_from = median_RESULTAT)  |> 
  mutate(
    factor_oyster_to_mussel = mussel / oyster,
    factor_mussel_to_oyster = oyster / mussel
  )

# # Save conversion factors
# writexl::write_xlsx(x = factors_metals, path = "../inst/results/data_contam/metals/conversion_factors_metal.xlsx")
```

Compute results in "equivalent mussels"
```{r}
data_metals <- data_metals |> 
  left_join(factors_metals, by = "PARAMETRE_LIBELLE") |> 
  mutate(RESULTAT_raw = RESULTAT) |> 
  mutate(RESULTAT = case_when(
    SPECIES == "oyster" ~ RESULTAT * factor_oyster_to_mussel,
    SPECIES == "mussel" ~ RESULTAT
  ))
```

## Get unit of measurment

Conclusion : original unit in mg/kg dw (equivalent in µg/g)

- 1979-2019 : "mg.kg-1"
- 2020-2024 : "mg/(kg MS)" and "mg.kg-1" or only "mg/(kg MS)"
```{r}
# data_metals |>
#   distinct(YEAR, UNITE, ESTUARY, PARAMETRE_LIBELLE, espece) |>  View()

data_metals |> 
  filter(ESTUARY == "Gironde", PARAMETRE_LIBELLE == "Mercury") |> 
  ggplot() +
  aes(x = YEAR, y = RESULTAT, colour = UNITE) +
  geom_point()
# => seems to be in the same range of values
```

```{r}
data_metals <- data_metals |> 
  mutate(UNITE = "µg/g dw") |> 
  # Get only one value per year
  group_by(ESTUARY, YEAR, PARAMETRE_LIBELLE) |>
  mutate(RESULTAT = mean(RESULTAT)) |> 
  ungroup() |> 
  # Round result values
  mutate(RESULTAT = round(RESULTAT, digits = 3))
```

Transform all results in µg/g ww (same as µg/kg):
```{r}
data_metals_ww <- data_metals |>
  left_join(esteem.overview::data_ROCCHMV_bio, by = "ESTUARY") |> 
  # transformation into ww concentrations
  mutate(RESULTAT = RESULTAT * (`Matière sèche_%` / 100)) |>  
  # Change to unified unit
  mutate(UNITE = "µg/g ww") |> 
  # Get only one value per year
  group_by(ESTUARY, YEAR, PARAMETRE_LIBELLE)  |>
  mutate(RESULTAT = mean(RESULTAT)) |> 
  ungroup() |>
  # Round result values
  mutate(RESULTAT = round(RESULTAT, digits = 3))
```

## Identify first and last periods

Identify the first and last 5 years of measurement
```{r}
# Compute minimal and maximal year of measurement by metal and estuary
min_max_year <- data_metals_ww |> 
  group_by(PARAMETRE_LIBELLE, ESTUARY) |> 
  summarise(min_year = min(YEAR),
            max_year = max(YEAR), .groups = "drop")

# Identify the first and last 5 years of measurement + last 15 years
data_metals_ww <- data_metals_ww |> 
  left_join(min_max_year, by = c("PARAMETRE_LIBELLE", "ESTUARY"))  |> 
  mutate(periode = case_when(
    YEAR <= min_year + 5 ~ "First 5 years",
    YEAR >= max_year - 5 ~ "Last 5 years",
    TRUE ~ NA
  )) |> 
  mutate(last_trend = case_when(
    YEAR >= max_year - 15 ~ "Last 15 years",
    TRUE ~ NA
  ))
```

# Time series graphs

## Last 15 years linear regression models

```{r}
res_metals_trends <- get_last_trend_info(data = data_metals_ww, norm = metals)

writexl::write_xlsx(res_metals_trends, "../inst/results/data_contam/metals/recent_trends_metals_µg_ww.xlsx")
```


## For metals with norm values
```{r}
res_metals_norm_trends <- res_metals_trends |> 
  filter(PARAMETRE_LIBELLE %in% c("Mercury", "Cadmium", "Lead"))

ggplot_metals_norm_ww <- data_metals_ww |> 
  filter(PARAMETRE_LIBELLE %in% c("Mercury", "Cadmium", "Lead")) |> 
  ggplot() +
  aes(x = YEAR,
      y = log(RESULTAT),
      colour = ESTUARY) +
  geom_line() +
  geom_smooth(se = FALSE) +
    theme(legend.position = "none",      # no legend
        axis.text.x = element_blank()) + # no x labels
  facet_grid(rows = vars(PARAMETRE_LIBELLE), scales = "free_y") +
  labs(colour = "Estuary", y = "Contamination level in log µg/g ww", x = NULL) +
# Add arrows for significant last trends (15 years)
  geom_text(
    data = res_metals_norm_trends,
    aes(x = last_year+1, y = log(last_resultat), label = symbol), size = rel(6),
    family = "DejaVu Sans",   # specify the font that contains arrows
    show.legend = FALSE) +
# Add environmental threshold when existing
  geom_text(
    data = res_metals_norm_trends,  
    aes(x = max(last_year)+1, 
        y = log(EC_MPC_µg_gww) - 0.2, 
        label = sprintf("%.2f", log(EC_MPC_µg_gww))),
    color = "red", size = 2.5
  ) +
  geom_hline(
    data = res_metals_norm_trends,
    aes(yintercept = log(EC_MPC_µg_gww)),
    color = "red", linetype = "dashed"
  )

ggplot_metals_norm_ww

ggsave(plot = ggplot_metals_norm_ww, "../inst/results/data_contam/metals/ggplot_metals_norm_ww_time_series.jpg",
       width = 10, height = 8, units = "cm")
```

## For the others
```{r}
res_metals_trends_2 <- res_metals_trends |> 
  filter(PARAMETRE_LIBELLE %in% c("Copper"))

ggplot_metals_ww <- data_metals_ww |> 
  filter(PARAMETRE_LIBELLE %in% c("Copper")) |> 
  ggplot() +
  aes(x = YEAR,
      y = log(RESULTAT),
      colour = ESTUARY) +
  geom_line() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom") +
  facet_grid(rows = vars(PARAMETRE_LIBELLE), scales = "free_y") +
  labs(colour = "Estuary", y = "Concentration \n in log µg/g ww", x = NULL) +
# Add arrows for significant last trends (15 years)
  geom_text(
    data = res_metals_trends_2,
    aes(x = last_year+1, y = log(last_resultat), label = symbol), size = rel(6),
    family = "DejaVu Sans",   # specify the font that contains arrows
    show.legend = FALSE)

ggplot_metals_ww

ggsave(plot = ggplot_metals_ww, "../inst/results/data_contam/metals/ggplot_metals_ww_time_series.jpg",
       width = 10, height = 4.7, units = "cm")
```

# First vs last period graphs

## For metals with norm values

```{r}
data_metals_norm_relative <- data_metals_ww |> 
  filter(PARAMETRE_LIBELLE %in% c("Mercury", "Cadmium", "Lead")) |> 
  left_join(metals) |> 
  mutate(RESULTAT = RESULTAT / EC_MPC_µg_gww * 100)

res_kruskal_metals_norm_relative <- get_kruskal_periods_results(
  data_with_periode = data_metals_norm_relative, 
  chemicals = c("Mercury", "Cadmium", "Lead"))
```

```{r}
plot_metals_norm_periode <- data_metals_norm_relative |> 
  filter(PARAMETRE_LIBELLE %in% c("Mercury", "Cadmium", "Lead")) |>
  drop_na(periode) |> # keep only data from first and last 5-years periods
  ggplot() +
  aes(x = ESTUARY, y = RESULTAT, colour = periode) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.9)) +
  scale_color_manual(values = c("purple", "deeppink")) +
  facet_grid(rows = vars(PARAMETRE_LIBELLE), scales = "free_y") +
  labs(x = NULL, y = "Deviation from limit value (% µg/g ww)", colour = "Period") +
  theme(legend.position = "none",      # no legend
        axis.text.x = element_blank()) + # no x labels
# Add long term trends if significant
  geom_text(data = res_kruskal_metals_norm_relative, 
            inherit.aes = FALSE, 
            aes(x = ESTUARY, y = 0.9*y_pos, label = trend), size = rel(6),
    family = "DejaVu Sans",   # specify the font that contains arrows
    show.legend = FALSE) 

plot_metals_norm_periode

ggsave(plot = plot_metals_norm_periode, "../inst/results/data_contam/metals/ggplot_metals_norm_periods.jpg", 
       width = 10, height = 8, units = "cm")
```

## For the others
```{r}
res_kruskal_metals <- get_kruskal_periods_results(
  data_with_periode = data_metals_ww |> filter(PARAMETRE_LIBELLE %in% c("Copper")), 
  chemicals = c("Copper"))
```

```{r}
plot_metals_periode <- data_metals_ww |> 
  filter(PARAMETRE_LIBELLE %in% c("Copper")) |>
  drop_na(periode) |> # keep only data from first and last 5-years periods
  ggplot() +
  aes(x = ESTUARY, y = RESULTAT, colour = periode) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.9)) +
  scale_color_manual(values = c("purple", "deeppink")) +
  facet_grid(rows = vars(PARAMETRE_LIBELLE), scales = "free_y") +
  labs(x = NULL, y = "Concentration \n in µg/g ww", colour = "Period") +
  theme(legend.position = "bottom") +
# Add long term trends if significant
  geom_text(data = res_kruskal_metals, 
            inherit.aes = FALSE, 
            aes(x = ESTUARY, y = 0.9*y_pos, label = trend), size = rel(6),
    family = "DejaVu Sans",   # specify the font that contains arrows
    show.legend = FALSE)

plot_metals_periode

ggsave(plot = plot_metals_periode, "../inst/results/data_contam/metals/ggplot_metals_periods.jpg", 
       width = 10, height = 4.7, units = "cm")
```

# General levels of contamination
```{r}
data_metals_ww_general_levels_estuary <- data_metals_ww |> 
  drop_na(periode) |> 
  group_by(ESTUARY, PARAMETRE_LIBELLE, periode) |> 
  summarise(median_ng_gww = median(RESULTAT), .groups = "drop") |> 
  arrange(PARAMETRE_LIBELLE) |> 
  pivot_wider(names_from = periode, values_from = median_ng_gww)

data_metals_ww_general_levels <- data_metals_ww |> 
  drop_na(periode) |> 
  group_by(PARAMETRE_LIBELLE, periode) |> 
  summarise(median_ng_gww = median(RESULTAT), .groups = "drop") |> 
  arrange(PARAMETRE_LIBELLE) |> 
  pivot_wider(names_from = periode, values_from = median_ng_gww)


writexl::write_xlsx(x = list(data_metals_ww_general_levels, data_metals_ww_general_levels_estuary), path = "../inst/results/data_contam/metals/summarised_levels_data_metals_ww.xlsx")
```


# General evaluation
```{r}
get_general_evaluation(
  family = "metals",
  suffix = "_µg_ww",
  res_family_trends = res_metals_norm_trends,
  res_kruskal = res_kruskal_metals_norm_relative)
```
