---
title: "data_contamination_PAH"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{data_contamination_PAH}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
library(esteem.overview)
library(segmented)
library(tidyverse)
library(here)
library(writexl)
```

# Get raw data for Polycyclic Aromatic Hydrocarbon (PAH) => thresholds in ng/g dw and ng/g ww
```{r}
PAH <- data.frame(PARAMETRE_LIBELLE = c("Fluoranthène", "Benzo(a)pyrène", "Anthracène", 
                                        "Naphtalène", "Phénanthrène", "Pyrène",
                                        "Benzo(a)anthracène",  "Benzo(g,h,i)pérylène"),
                   EAC_OSPAR_ng_gdw = c(110, 600, 290, 340, 1700, 100, 80, 110),
                   EQS_VGE_ng_gww = c(30, 5, 47.47, 19.7, NA, NA, NA, NA)
)

data_PAH <- esteem.overview::data_ROCCHMV_cleaned |> ungroup() |> 
  filter(support == "Bivalve") |> 
  select(-c(SUPPORT_NIVEAU_PRELEVEMENT, support, niveau)) |> 
  filter(PARAMETRE_LIBELLE %in% PAH$PARAMETRE_LIBELLE) |> 
  left_join(esteem.overview::data_ROCCHMV_bio, by = "ESTUARY")
```

## Get the same unit of measurment

Different units:

- 1994-2016: µg.kg-1				
- 2018-2024: µg.kg-1, p.h. (eq ng.g-1, p.h.)
```{r}
data_PAH |> 
  distinct(UNITE)

data_PAH |> 
  mutate(YEAR = as.factor(YEAR)) |> 
  distinct(YEAR, UNITE, ESTUARY, PARAMETRE_LIBELLE) |>  View(title = "PAH units")

# # Check difference dw ww for a common YEAR x ESTUARY x chemical
# data_PAH |> 
#   filter(YEAR == 2020, ESTUARY == "Loire", PARAMETRE_LIBELLE == "Benzo(a)anthracène")
```

Transform all results in ng/g dw (same as µg/kg):
```{r}
data_PAH_dw <- data_PAH |>
  # transformation into dw concentrations
  mutate(RESULTAT = case_when(
    UNITE == "µg.kg-1, p.h." ~ RESULTAT / (`Matière sèche_%` / 100),
    UNITE == "µg.kg-1" ~ RESULTAT
  )) |>  
  # Change to unified unit
  mutate(UNITE = "ng/g dw") |> 
  # Get only one value per year
  group_by(ESTUARY, YEAR, PARAMETRE_LIBELLE) |>
  # Round result values
  mutate(RESULTAT = round(RESULTAT, digits = 3))
```

Transform all results in ng/g ww (same as µg/kg):
```{r}
data_PAH_ww <- data_PAH |>
  # transformation into ww concentrations
  mutate(RESULTAT = case_when(
    UNITE == "µg.kg-1, p.h." ~ RESULTAT,
    UNITE == "µg.kg-1" ~ RESULTAT * (`Matière sèche_%` / 100)
  )) |>  
  # Change to unified unit
  mutate(UNITE = "ng/g ww") |> 
  # Get only one value per year
  group_by(ESTUARY, YEAR, PARAMETRE_LIBELLE) |>
  # Round result values
  mutate(RESULTAT = round(RESULTAT, digits = 3))
```

## Importance of the species matrix

- Gironde : only C. gigas
- Seine : only M. edulis
- Loire : 
  - 1994-2021: M. edulis
  - 2022-2024: M. edulis + M.galloprovincialis
  - 2018, 2021, 2024: Crassostrea gigas
```{r}
data_PAH_dw |> 
  distinct(ESTUARY, espece)

data_PAH_dw |> 
  filter(ESTUARY == "Loire") |> 
  distinct(ESTUARY, YEAR, espece)
```

No species effect
```{r}
data_PAH_dw |> 
  filter(ESTUARY == "Loire", YEAR > 2010) |>
  ggplot() +
  aes(x = YEAR, y = RESULTAT, colour = espece) +
  geom_line() +
  facet_wrap(vars(PARAMETRE_LIBELLE), scales = "free_y")
```

Simplify with two species groups
```{r}
data_PAH_dw <- data_PAH_dw |> 
    mutate(SPECIES = case_when(
    espece == "Crassostrea gigas (huître creuse)" ~ "oyster",
    espece %in% c("Mytilus edulis (moule commune)",
                  "Mytilus edulis + galloprovincialis (moule)") ~ "mussel"
  ))

data_PAH_ww <- data_PAH_ww |> 
    mutate(SPECIES = case_when(
    espece == "Crassostrea gigas (huître creuse)" ~ "oyster",
    espece %in% c("Mytilus edulis (moule commune)",
                  "Mytilus edulis + galloprovincialis (moule)") ~ "mussel"
  ))
```

## Identify first and last periods

Identify the first and last 5 years of measurement
```{r}
# Compute minimal and maximal year of measurement by compounds and estuary
min_max_year <- data_PAH_dw |> 
  group_by(PARAMETRE_LIBELLE, ESTUARY) |> 
  summarise(min_year = min(YEAR),
            max_year = max(YEAR), .groups = "drop")

# Identify the first and last 5 years of measurement
data_PAH_dw <- data_PAH_dw |> 
  left_join(min_max_year, by = c("PARAMETRE_LIBELLE", "ESTUARY"))  |> 
  mutate(periode = case_when(
    YEAR <= min_year + 5 ~ "First 5 years",
    YEAR >= max_year - 5 ~ "Last 5 years",
    TRUE ~ NA
  ))
```

```{r}
# Compute minimal and maximal year of measurement by compounds and estuary
min_max_year <- data_PAH_ww |> 
  group_by(PARAMETRE_LIBELLE, ESTUARY) |> 
  summarise(min_year = min(YEAR),
            max_year = max(YEAR), .groups = "drop")

# Identify the first and last 5 years of measurement
data_PAH_ww <- data_PAH_ww |> 
  left_join(min_max_year, by = c("PARAMETRE_LIBELLE", "ESTUARY"))  |> 
  mutate(periode = case_when(
    YEAR <= min_year + 5 ~ "First 5 years",
    YEAR >= max_year - 5 ~ "Last 5 years",
    TRUE ~ NA
  ))
```

# Time series graphs
```{r}
data_PAH_dw |>  
ggplot() +
  aes(x = YEAR, y = RESULTAT, color = ESTUARY) +
  geom_line() +
  facet_grid(vars(PARAMETRE_LIBELLE), scales = "free_y") +
  labs(x = NULL)
```

```{r}
PAH_names <- PAH$PARAMETRE_LIBELLE 
estuary_names <- c("Gironde", "Loire", "Seine")
params <- expand.grid(estuary_names, PAH_names)
params_ww <- expand.grid(estuary_names, PAH_names[1:4])

# in dw
fun_seg_model_PAH_dw <- function(chemical, estuary){
  fun_seg_model(data = data_PAH_dw,
                folder = "PAH",
                chemical = chemical, 
                estuary = estuary)
}
res_PAH_dw_trends <- pmap_dfr(params, ~ fun_seg_model_PAH_dw(estuary = ..1, chemical = ..2))
res_PAH_dw_trends <- res_PAH_dw_trends |> 
  left_join(PAH, by = "PARAMETRE_LIBELLE")

# in ww
fun_seg_model_PAH_ww <- function(chemical, estuary){
  fun_seg_model(data = data_PAH_ww,
                folder = "PAH",
                chemical = chemical, 
                estuary = estuary)
}
res_PAH_ww_trends <- pmap_dfr(params_ww, ~ fun_seg_model_PAH_ww(estuary = ..1, chemical = ..2))
res_PAH_ww_trends <- res_PAH_ww_trends |> 
  left_join(PAH, by = "PARAMETRE_LIBELLE")
```

```{r}
ggplot_PAH_dw <- ggplot(data_PAH_dw) +
  aes(x = YEAR,
      y = log(RESULTAT),
      colour = ESTUARY) +
  geom_line() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom") +
  facet_grid(rows = vars(
    fct_relevel(PARAMETRE_LIBELLE, PAH_names)
  ), scales = "free_y") +
  labs(colour = "Estuary", y = "Contamination level in log ng/g dw", x = NULL) +
# Add last break dates of segmented models and arrows for significant last trends
  geom_vline(data = res_PAH_dw_trends, aes(xintercept = break_dates, colour = ESTUARY)) +
  geom_text(
    data = res_PAH_dw_trends,
    aes(x = last_year+1, y = log(last_resultat), label = symbol), size = rel(6),
    family = "DejaVu Sans",   # specify the font that contains arrows
    show.legend = FALSE) +
# Add environmental threshold when existing => EAC OSPAR 2023
  geom_text(
    data = res_PAH_dw_trends,  
    aes(x = last_year+1, y = max(log(last_resultat))*1.1, label = sprintf("%.2f", log(EAC_OSPAR_ng_gdw))),
    color = "red", size = 2.5
  ) +
  geom_hline(
    data = res_PAH_dw_trends,
    aes(yintercept = log(EAC_OSPAR_ng_gdw)),
    color = "red", linetype = "dashed"
  )

ggplot_PAH_dw

ggsave(plot = ggplot_PAH_dw, "../inst/results/data_contam/PAH/ggplot_PAH_dw_time_series.jpg",
       width = 10, height = 15, units = "cm")
```

```{r}
ggplot_PAH_ww <- data_PAH_ww |> 
  filter(PARAMETRE_LIBELLE %in% PAH$PARAMETRE_LIBELLE[1:4])|> 
  ggplot() +
  aes(x = YEAR,
      y = log(RESULTAT),
      colour = ESTUARY) +
  geom_line() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom") +
  facet_grid(rows = vars(
    fct_relevel(PARAMETRE_LIBELLE, PAH_names)
  ), scales = "free_y") +
  labs(colour = "Estuary", y = "Contamination level in log ng/g ww", x = NULL) +
# Add last break dates of segmented models and arrows for significant last trends
  geom_vline(data = res_PAH_ww_trends |> filter(PARAMETRE_LIBELLE %in% PAH$PARAMETRE_LIBELLE[1:4]), 
             aes(xintercept = break_dates, colour = ESTUARY)) +
  geom_text(
    data = res_PAH_ww_trends |> filter(PARAMETRE_LIBELLE %in% PAH$PARAMETRE_LIBELLE[1:4]),
    aes(x = last_year+1, y = log(last_resultat), label = symbol), size = rel(6),
    family = "DejaVu Sans",   # specify the font that contains arrows
    show.legend = FALSE) +
# Add environmental threshold when existing => EAC OSPAR 2023
  geom_text(
    data = res_PAH_ww_trends |> filter(PARAMETRE_LIBELLE %in% PAH$PARAMETRE_LIBELLE[1:4]),  
    aes(x = last_year+1, y = max(log(last_resultat))*1.1, label = sprintf("%.2f", log(EQS_VGE_ng_gww))),
    color = "red", size = 2.5
  ) +
  geom_hline(
    data = res_PAH_ww_trends |> filter(PARAMETRE_LIBELLE %in% PAH$PARAMETRE_LIBELLE[1:4]),
    aes(yintercept = log(EQS_VGE_ng_gww)),
    color = "red", linetype = "dashed"
  )

ggplot_PAH_ww

ggsave(plot = ggplot_PAH_ww, "../inst/results/data_contam/PAH/ggplot_PAH_ww_time_series.jpg",
       width = 10, height = 10, units = "cm")
```

# First vs last period graphs

In dw
```{r}
data_PAH_dw_norm_relative <- data_PAH_dw |> 
  left_join(PAH) |> 
  mutate(RESULTAT = RESULTAT / EAC_OSPAR_ng_gdw * 100)

res_kruskal_PAH_dw <- get_kruskal_periods_results(data_with_periode = data_PAH_dw_norm_relative, 
                             chemicals = PAH$PARAMETRE_LIBELLE)
```

```{r}
plot_PAH_dw_periode <- data_PAH_dw_norm_relative |> 
  drop_na(periode) |> # keep only data from first and last 5-years periods
  ggplot() +
  aes(x = ESTUARY, y = RESULTAT, colour = periode) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.9)) +
  scale_color_manual(values = c("purple", "deeppink")) +
  facet_grid(rows = vars(fct_relevel(PARAMETRE_LIBELLE, PAH$PARAMETRE_LIBELLE)), scales = "free_y") +
  labs(x = NULL, y = "Deviation from limit value (% ng/g dw)", colour = "Period") +
  theme(legend.position = "bottom") +
# Add long term trends if significant
  geom_text(data = res_kruskal_PAH_dw, 
            inherit.aes = FALSE, 
            aes(x = ESTUARY, y = 0.9*y_pos, label = trend), size = rel(6),
    family = "DejaVu Sans",   # specify the font that contains arrows
    show.legend = FALSE)

plot_PAH_dw_periode

ggsave(plot = plot_PAH_dw_periode, "../inst/results/data_contam/PAH/ggplot_PAH_dw_periods.jpg", 
       width = 10, height = 15, units = "cm")
```

In ww

```{r}
data_PAH_ww_norm_relative <- data_PAH_ww |> 
  left_join(PAH) |> 
  mutate(RESULTAT = RESULTAT / EQS_VGE_ng_gww * 100) |> 
  drop_na(RESULTAT)

res_kruskal_PAH_ww <- get_kruskal_periods_results(data_with_periode = data_PAH_ww_norm_relative, 
                             chemicals = PAH$PARAMETRE_LIBELLE[1:4])
```

```{r}
plot_PAH_ww_periode <- data_PAH_ww_norm_relative |> 
  drop_na(periode) |> # keep only data from first and last 5-years periods
  ggplot() +
  aes(x = ESTUARY, y = RESULTAT, colour = periode) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.9)) +
  scale_color_manual(values = c("purple", "deeppink")) +
  facet_grid(rows = vars(fct_relevel(PARAMETRE_LIBELLE, PAH$PARAMETRE_LIBELLE)), scales = "free_y") +
  labs(x = NULL, y = "Deviation from limit value (% ng/g ww)", colour = "Period") +
  theme(legend.position = "bottom") +
# Add long term trends if significant
  geom_text(data = res_kruskal_PAH_ww, 
            inherit.aes = FALSE, 
            aes(x = ESTUARY, y = 0.9*y_pos, label = trend), size = rel(6),
    family = "DejaVu Sans",   # specify the font that contains arrows
    show.legend = FALSE)

plot_PAH_ww_periode

ggsave(plot = plot_PAH_ww_periode, "../inst/results/data_contam/PAH/ggplot_PAH_ww_periods.jpg", 
       width = 10, height = 10, units = "cm")
```

# General evaluation

In ww
```{r}
get_general_evaluation(family = "PAH",
                       suffix = "_ww",
                       res_family_trends = res_PAH_ww_trends,
                       res_kruskal = res_kruskal_PAH_ww)
```

In dw
```{r}
get_general_evaluation(family = "PAH",
                       suffix = "_dw",
                       res_family_trends = res_PAH_dw_trends,
                       res_kruskal = res_kruskal_PAH_dw)
```
