---
title: "data_contamination_PFOS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{data_contamination_PFOS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(esteem.overview)
library(tidyverse)
library(here)
```

# Get raw data for Polychlorinated biphenyls indicators (PFOS) => thresholds in ng/g lw
```{r}
PFOS <- data.frame(PARAMETRE_LIBELLE = "PFOS",
                   EQS_EC_ng_gww = 9.1)

data_PFOS <- esteem.overview::data_ROCCHMV_cleaned |> ungroup() |> 
  filter(support == "Bivalve") |> 
  select(-c(SUPPORT_NIVEAU_PRELEVEMENT, support, niveau)) |> 
  filter(PARAMETRE_LIBELLE %in% PFOS$PARAMETRE_LIBELLE) |> 
  left_join(esteem.overview::data_ROCCHMV_bio, by = "ESTUARY")

```

## Get the same unit of measurment

Different units:

- 2010-2017: ng.g-1				
- 2018-2024: µg.kg-1, p.h.				
- 2021 Seine: ng.kg-1, p.h.
```{r}
data_PFOS |> 
  distinct(UNITE)

# data_PFOS |> 
#   mutate(YEAR = as.factor(YEAR)) |> 
#   distinct(YEAR, UNITE, ESTUARY, PARAMETRE_LIBELLE) |>  View(title = "PFOS units")
```

Transform all results in ng/g ww (same as µg/kg):
```{r}
data_PFOS_ww <- data_PFOS |>
  # transformation into ww concentrations
  mutate(RESULTAT = case_when(
    UNITE == "µg.kg-1, p.h." ~ RESULTAT,
    UNITE == "ng.kg-1, p.h." ~ RESULTAT / 1000,
    UNITE == "ng.g-1" ~ RESULTAT * (`Matière sèche_%` / 100),
  )) |>  
  # Change to unified unit
  mutate(UNITE = "ng/g ww") |> 
  # Get only one value per year
  group_by(ESTUARY, YEAR, PARAMETRE_LIBELLE) |>
  mutate(RESULTAT = mean(RESULTAT)) |> 
  ungroup() |> 
  # Round result values
  mutate(RESULTAT = round(RESULTAT, digits = 3))
```

```{r}
data_PFOS_ww |> 
  group_by(ESTUARY, PARAMETRE_LIBELLE) |> 
  summarise(mean = mean(RESULTAT))
```


## Importance of the species matrix

- Gironde : only C. gigas
- Seine : only M. edulis
- Loire : 
  - 2010-2018: M. edulis
  - 2023-2024: M. edulis + M.galloprovincialis
  - 2024: C. gigas
```{r}
data_PFOS_ww |> 
  distinct(ESTUARY, espece)

data_PFOS_ww |> 
  filter(ESTUARY == "Loire") |> 
  distinct(ESTUARY, YEAR, espece)
```

```{r}
data_PFOS_ww |> 
  filter(ESTUARY == "Loire") |>
  ggplot() +
  aes(x = YEAR, y = RESULTAT, colour = espece) +
  geom_line() +
  facet_wrap(vars(PARAMETRE_LIBELLE), scales = "free_y")
```

Simplify with two species groups
```{r}
data_PFOS_ww <- data_PFOS_ww |> 
    mutate(SPECIES = case_when(
    espece == "Crassostrea gigas (huître creuse)" ~ "oyster",
    espece %in% c("Mytilus edulis (moule commune)",
                  "Mytilus edulis + galloprovincialis (moule)") ~ "mussel"
  ))
```

# Time series graphs

## Get the monitoring periods
```{r}
data_PFOS_ww |> 
  group_by(ESTUARY, PARAMETRE_LIBELLE) |> 
  summarise(min_year = min(YEAR),
            max_year = max(YEAR))
``` 

## Spearman's correlation coefficient + test

```{r}
# Compute for all time serie
data_PFOS_ww <- data_PFOS_ww |> 
  mutate(last_trend = "Last 15 years")

res_PFOS_ww_trends <- get_last_trend_info(data = data_PFOS_ww, norm = PFOS)

writexl::write_xlsx(res_PFOS_ww_trends, "../inst/results/data_contam/PFOS/recent_trends_PFOS_ww.xlsx")
```

```{r}
ggplot_PFOS_ww <- ggplot(data_PFOS_ww) +
  aes(x = YEAR,
      y = log(RESULTAT),
      colour = ESTUARY) +
  geom_line() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom") +
  facet_grid(rows = vars(
    fct_relevel(PARAMETRE_LIBELLE, PFOS$PARAMETRE_LIBELLE)
  ), scales = "free_y") +
  labs(colour = "Estuary", y = "Contamination level \n in log ng/g ww", x = NULL) +
# Add arrows for significant last trends
  geom_text(
    data = res_PFOS_ww_trends,
    aes(x = max(last_year)+1, y = log(last_resultat), label = symbol), size = rel(6),
    family = "DejaVu Sans",   # specify the font that contains arrows
    show.legend = FALSE) +
# Add environmental threshold when existing
  geom_text(
    data = res_PFOS_ww_trends,  
    aes(x = max(last_year)+1, y = log(EQS_EC_ng_gww) +0.5, label = sprintf("%.2f", log(EQS_EC_ng_gww))),
    color = "red", size = 2.5
  ) +
  geom_hline(
    data = res_PFOS_ww_trends,
    aes(yintercept = log(EQS_EC_ng_gww)),
    color = "red", linetype = "dashed"
  )

ggplot_PFOS_ww

ggsave(plot = ggplot_PFOS_ww, "../inst/results/data_contam/PFOS/ggplot_PFOS_ww_time_series.jpg",
       width = 10, height = 7, units = "cm")
```

# First vs last period graphs
```{r}
data_PFOS_ww_norm_relative <- data_PFOS_ww |> 
  left_join(PFOS) |> 
  mutate(RESULTAT = RESULTAT / EQS_EC_ng_gww * 100)
```

```{r}
plot_PFOS_ww_periode <- data_PFOS_ww_norm_relative |> 
  ggplot() +
  aes(x = ESTUARY, y = RESULTAT) +
  geom_boxplot(width = 0.6, colour = "deeppink") +
  facet_grid(rows = vars(PARAMETRE_LIBELLE), scales = "free_y") +
  labs(x = NULL, y = "Deviation from limit value \n (% ng/g ww)", colour = "Period") +
  theme(legend.position = "bottom")

plot_PFOS_ww_periode

ggsave(plot = plot_PFOS_ww_periode, "../inst/results/data_contam/PFOS/ggplot_PFOS_ww_periods.jpg", 
       width = 10, height = 7, units = "cm")
```

## General levels of contamination
```{r}
data_PFOS_ww_levels <- data_PFOS_ww |> 
  group_by(ESTUARY, PARAMETRE_LIBELLE) |> 
  summarise(median_ng_gww = median(RESULTAT), .groups = "drop") |> 
  arrange(PARAMETRE_LIBELLE)

data_PFOS_ww_normrelative <- data_PFOS_ww_norm_relative |> 
  group_by(ESTUARY, PARAMETRE_LIBELLE) |> 
  summarise(median_percent = median(RESULTAT), .groups = "drop") |> 
  arrange(PARAMETRE_LIBELLE) |>  
  select(ESTUARY, PARAMETRE_LIBELLE, median_percent)

data_PFOS_general_levels <- full_join(data_PFOS_ww_levels, data_PFOS_ww_normrelative)

writexl::write_xlsx(x = data_PFOS_general_levels, path = "../inst/results/data_contam/PFOS/summarised_levels_data_PFOS_ww.xlsx")
```


# General evaluation
```{r}
# trends
res_PFOS_ww_trends

# color dot from percentage
# blue: [0-50%[, green: [50-100%[, yellow:[100-150%[, orange: [150-300%[, red: [300-500%[, black: > 500%
data_PFOS_ww_norm_relative |> 
  group_by(ESTUARY, PARAMETRE_LIBELLE) |> 
  summarise(median = median(RESULTAT))
```

