---
title: "data_contamination_brominated"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{data_contamination_brominated}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
library(esteem.overview)
library(segmented)
library(tidyverse)
library(here)
```

# Get raw data for brominated flame retardand compounds => thresholds in ng/g ww
```{r}
data <- esteem.overview::data_ROCCHMV_cleaned |> ungroup() |> 
  filter(support == "Bivalve") |> 
  select(-c(SUPPORT_NIVEAU_PRELEVEMENT, support, niveau))
```

Hexabromocyclododecane isomers (HBCDD)
```{r}
HBCDD_names <- c("Alpha-HBCDD", "Beta-HBCDD", "Gamma-HBCDD")
data_HBCDD <- data |> filter(PARAMETRE_LIBELLE %in% HBCDD_names)
```

Total sum PBDE congeners
```{r}
PBDE_names <- c("PBDE 28", "PBDE 47", "PBDE 99", "PBDE 100", "PBDE 153", "PBDE 154")
data_PBDE <- data |> filter(PARAMETRE_LIBELLE %in% PBDE_names) 
```

Brominated flame retardant
```{r}
data_brominated <- full_join(data_HBCDD, data_PBDE)  |> 
  left_join(esteem.overview::data_ROCCHMV_bio, by = "ESTUARY")
```

## Get the same unit of measurment

Different units:

- 2008-2014: ng.kg-1
- 2015-2016: ng.kg-1 & µg/(kg MS)
- 2017: ng.g-1, p.h.& ng.kg-1
- 2018-2024: ng.g-1, p.h.
```{r}
data_brominated |> 
  distinct(UNITE)

data_brominated |> 
  mutate(YEAR = as.factor(YEAR)) |> 
  distinct(YEAR, UNITE, ESTUARY, PARAMETRE_LIBELLE) |>  View(title = "Brominated units")
```

Transform all results in ng/g ww (same as µg/kg):
```{r}
data_brominated_ww <- data_brominated |>
  # transformation into dw concentrations
  mutate(RESULTAT = case_when(
    UNITE == "ng.g-1, p.h." ~ RESULTAT,
    UNITE == "ng.kg-1" ~ RESULTAT / 1000 * (`Matière sèche_%` / 100) ,
    UNITE == "µg/(kg MS)" ~ RESULTAT * (`Matière sèche_%` / 100)
  )) |>  
  # Change to unified unit
  mutate(UNITE = "ng/g ww") |> 
  # Get only one value per year
  group_by(ESTUARY, YEAR, PARAMETRE_LIBELLE) |> 
  mutate(RESULTAT = mean(RESULTAT)) |> 
  ungroup() |> 
  # Round result values
  mutate(RESULTAT = round(RESULTAT, digits = 4))

# # Checking
# data_brominated |> filter(YEAR %in% c(2015, 2017,2019, 2021), ESTUARY == "Gironde", PARAMETRE_LIBELLE == "sum_DDT")
# data_brominated_ww |> filter(YEAR %in% c(2015, 2017,2019, 2021), ESTUARY == "Gironde", PARAMETRE_LIBELLE == "sum_DDT")
```

## Importance of the species matrix

- Gironde : only C. gigas
- Seine : only M. edulis
- Loire : 
  - 2008-2021: M. edulis
  - 2022-2024: M. edulis + M.galloprovincialis
```{r}
data_brominated_ww |> 
  distinct(ESTUARY, espece)

data_brominated_ww |> 
  filter(ESTUARY == "Loire") |> 
  distinct(ESTUARY, YEAR, espece)
```

No comparison possible between oysters and mussels (no oysters in Loire)
Strong difference possible for Beta-HBCDD for M. galloprovincialis but no parallel measurments
```{r}
data_brominated_ww |> 
  filter(ESTUARY == "Loire") |>
  ggplot() +
  aes(x = YEAR, y = RESULTAT, colour = espece) +
  geom_line() +
  facet_wrap(vars(PARAMETRE_LIBELLE), scales = "free_y")
```

Simplify with two species groups
```{r}
data_brominated_ww <- data_brominated_ww |> 
    mutate(SPECIES = case_when(
    espece == "Crassostrea gigas (huître creuse)" ~ "oyster",
    espece %in% c("Mytilus edulis (moule commune)",
                  "Mytilus edulis + galloprovincialis (moule)") ~ "mussel"
  ))
```

## Compute the sum of PBDE isomers
```{r}
data_PBDE_ww <- data_brominated_ww |> 
  filter(PARAMETRE_LIBELLE %in% PBDE_names) |>
  group_by(ESTUARY, YEAR, espece, UNITE) |>
  summarise(
    RESULTAT = sum(RESULTAT),
    .groups = "drop") |>
  mutate(PARAMETRE_LIBELLE = "sum_PBDE")
  
data_HBCDD_ww <- data_brominated_ww |> 
  filter(PARAMETRE_LIBELLE %in% HBCDD_names) |> 
  select(ESTUARY, YEAR, espece, UNITE, PARAMETRE_LIBELLE, RESULTAT)

data_brominated_ww <- full_join(data_PBDE_ww, data_HBCDD_ww) |> 
  left_join(esteem.overview::data_ROCCHMV_bio, by = "ESTUARY")

brominated_names <- c(HBCDD_names, "sum_PBDE")
brominated <- data.frame(PARAMETRE_LIBELLE = brominated_names,
                         EQS_EC_ng_gww = c(167, 167, 167, 0.0085))
```

## Identify first and last periods

Identify the first and last 3 years of measurement
```{r}
# Compute minimal and maximal year of measurement by compounds and estuary
min_max_year <- data_brominated_ww |> 
  group_by(PARAMETRE_LIBELLE, ESTUARY) |> 
  summarise(min_year = min(YEAR),
            max_year = max(YEAR), .groups = "drop")

# Identify the first and last 2 years of measurement
data_brominated_ww <- data_brominated_ww |> 
  left_join(min_max_year, by = c("PARAMETRE_LIBELLE", "ESTUARY"))  |> 
  mutate(periode = case_when(
    YEAR <= min_year + 2 ~ "First 2 years",
    YEAR >= max_year - 2 ~ "Last 2 years",
    TRUE ~ NA
  ))
```

# Time series graphs
```{r}
data_brominated_ww |>  
ggplot() +
  aes(x = YEAR, y = RESULTAT, color = ESTUARY) +
  geom_line() +
  facet_grid(vars(PARAMETRE_LIBELLE), scales = "free_y") +
  labs(x = NULL)
```

```{r}
fun_seg_model_brominated <- function(chemical, estuary, n_break){
  fun_seg_model(data = data_brominated_ww,
                folder = "brominated",
                chemical = chemical, 
                estuary = estuary, n_break = n_break)
}

estuary_names <- c("Gironde", "Loire", "Seine")
params <- expand.grid(estuary_names, brominated_names)

res_brominated_ww_trends <- pmap_dfr(params, ~ fun_seg_model_brominated(estuary = ..1, chemical = ..2))

res_brominated_ww_trends <- res_brominated_ww_trends |> 
  left_join(brominated, by = "PARAMETRE_LIBELLE")
```

```{r}
ggplot_brominated_ww <- ggplot(data_brominated_ww) +
  aes(x = YEAR,
      y = log(RESULTAT),
      colour = ESTUARY) +
  geom_line() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom") +
  facet_grid(vars(PARAMETRE_LIBELLE), scales = "free_y") +
  labs(colour = "Estuary", y = "Contamination level in ng/g ww", x = NULL) +
# Add last break dates of segmented models and arrows for significant last trends
  geom_vline(data = res_brominated_ww_trends, aes(xintercept = break_dates, colour = ESTUARY)) +
  geom_text(
    data = res_brominated_ww_trends,
    aes(x = last_year+1, y = log(last_resultat), label = symbol), size = rel(6),
    family = "DejaVu Sans",   # specify the font that contains arrows
    show.legend = FALSE)+
# Add environmental threshold when existing => UC QS(HH)
  geom_text(
    data = res_brominated_ww_trends,
    aes(x = last_year+1, y = log(EQS_EC_ng_gww)*1.3,
        label = sprintf("%.2f", log(EQS_EC_ng_gww))),
    color = "red", size = 2.5
  ) +
  geom_hline(
    data = res_brominated_ww_trends,
    aes(yintercept = log(EQS_EC_ng_gww)),
    color = "red", linetype = "dashed"
  )

ggplot_brominated_ww

ggsave(plot = ggplot_brominated_ww, "../inst/results/data_contam/brominated/ggplot_brominated_ww_time_series.jpg",
       width = 10, height = 10, units = "cm")
```

# First vs last period graphs
```{r}
data_brominated_ww_norm_relative <- data_brominated_ww |> 
  left_join(brominated) |> 
  mutate(RESULTAT = RESULTAT / EQS_EC_ng_gww * 100) |> 
  drop_na(RESULTAT)

res_kruskal_brominated <- get_kruskal_periods_results(
  data_with_periode = data_brominated_ww_norm_relative, 
  chemicals = brominated_names)
```

```{r}
plot_brominated_periode <- data_brominated_ww_norm_relative |> 
  drop_na(periode) |> # keep only data from first and last 5-years periods
  ggplot() +
  aes(x = ESTUARY, y = RESULTAT, colour = periode) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.9)) +
  scale_color_manual(values = c("purple", "deeppink")) +
  facet_grid(rows = vars(PARAMETRE_LIBELLE), scales = "free_y") +
  labs(x = NULL, y = "Concentration in ng/g ww", colour = "Period") +
  theme(legend.position = "bottom") +
# Add long term trends if significant
  geom_text(data = res_kruskal_brominated, 
            inherit.aes = FALSE, 
            aes(x = ESTUARY, y = 0.9*y_pos, label = trend), size = rel(6),
    family = "DejaVu Sans",   # specify the font that contains arrows
    show.legend = FALSE)

plot_brominated_periode

ggsave(plot = plot_brominated_periode, "../inst/results/data_contam/brominated/ggplot_brominated_ww_periods.jpg", 
       width = 10, height = 10, units = "cm")
```

# General evaluation
```{r}
get_general_evaluation(family = "brominated",
                       suffix = "_ww",
                       res_family_trends = res_brominated_ww_trends,
                       res_kruskal = res_kruskal_brominated)
```
