---
title: "data_contamination_OC_pesticides"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{data_contamination_OC_pesticides}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(esteem.overview)
library(tidyverse)
library(here)
```

# Get raw data for OC_pesticides compounds => thresholds in ng/g ww
```{r}
data <- esteem.overview::data_ROCCHMV_cleaned |> ungroup() |> 
  filter(support == "Bivalve") |> 
  select(-c(SUPPORT_NIVEAU_PRELEVEMENT, support, niveau))
```

Organochlorine pesticides: Lindane (gamma-HCH) and DDT isomers
```{r}
lindane <- c("Gamma-HCH")
data_lindane <- data |> filter(PARAMETRE_LIBELLE %in% lindane)

DDT <- c("p,p'-DDT", "o,p'-DDT", "p,p'-DDE", "p,p'-DDD")
data_DDT <- data |> filter(PARAMETRE_LIBELLE %in% DDT)

data_OC_pesticides <- full_join(data_lindane, data_DDT)  |> 
  left_join(esteem.overview::data_ROCCHMV_bio, by = "ESTUARY")
```

## Get the same unit of measurment

Different units:

- 1979-2016: µg.kg-1		
- 2017-2018 : µg/(kg MS)
- 2019-2020: ng.g-1, p.h. (eq µg.kg-1, p.h.)
- 2021-2023: µg.kg-1, p.h.
```{r}
data_OC_pesticides |> 
  distinct(UNITE)

# data_OC_pesticides |> 
#   mutate(YEAR = as.factor(YEAR)) |> 
#   distinct(YEAR, UNITE, ESTUARY, PARAMETRE_LIBELLE) |>  View(title = "OC_pesticides units")
```

Transform all results in ng/g ww (same as µg/kg):
```{r}
data_OC_pesticides_ww <- data_OC_pesticides |>
  # transformation into dw concentrations
  mutate(RESULTAT = case_when(
    UNITE == "ng.g-1, p.h." ~ RESULTAT,
    UNITE == "µg.kg-1, p.h." ~ RESULTAT,
    UNITE == "µg.kg-1" ~ RESULTAT * (`Matière sèche_%` / 100),
    UNITE == "µg/(kg MS)" ~ RESULTAT * (`Matière sèche_%` / 100)
  )) |>  
  # Change to unified unit
  mutate(UNITE = "ng/g ww") |> 
  # Get only one value per year
  group_by(ESTUARY, YEAR, PARAMETRE_LIBELLE) |> 
  mutate(RESULTAT = mean(RESULTAT)) |> 
  ungroup() |> 
  # Round result values
  mutate(RESULTAT = round(RESULTAT, digits = 4))

# # Checking
# data_OC_pesticides |> filter(YEAR %in% c(2015, 2017,2019, 2021), ESTUARY == "Gironde", PARAMETRE_LIBELLE == "Gamma-HCH")
# data_OC_pesticides_ww |> filter(YEAR %in% c(2015, 2017,2019, 2021), ESTUARY == "Gironde", PARAMETRE_LIBELLE == "Gamma-HCH")
```

## Importance of the species matrix

- Gironde : only C. gigas
- Seine : only M. edulis
- Loire : 
  - 1979-2021: M. edulis
  - 2022-2023: M. edulis + M.galloprovincialis
```{r}
data_OC_pesticides_ww |> 
  distinct(ESTUARY, espece)

data_OC_pesticides_ww |> 
  filter(ESTUARY == "Loire") |> 
  distinct(ESTUARY, YEAR, espece)
```

No comparison possible between oysters and musseles (no oysters in Loire):
```{r}
data_OC_pesticides_ww |> 
  filter(ESTUARY == "Loire", YEAR > 2000) |>
  ggplot() +
  aes(x = YEAR, y = RESULTAT, colour = espece) +
  geom_line() +
  facet_wrap(vars(PARAMETRE_LIBELLE), scales = "free_y") +
  theme_esteem()
```

Simplify with two species groups
```{r}
data_OC_pesticides_ww <- data_OC_pesticides_ww |> 
    mutate(SPECIES = case_when(
    espece == "Crassostrea gigas (huître creuse)" ~ "oyster",
    espece %in% c("Mytilus edulis (moule commune)",
                  "Mytilus edulis + galloprovincialis (moule)") ~ "mussel"
  ))
```

## Compute the sum of DDT

```{r}
data_DDT_ww <- data_OC_pesticides_ww |> 
  filter(PARAMETRE_LIBELLE %in% DDT) |>
  group_by(ESTUARY, YEAR, espece, UNITE) |>
  summarise(
    RESULTAT = sum(RESULTAT),
    .groups = "drop") |>
  mutate(PARAMETRE_LIBELLE = "sum_DDT")

data_lindane_ww <- data_OC_pesticides_ww |> 
  filter(PARAMETRE_LIBELLE %in% lindane)

data_OC_pesticides_ww <- full_join(data_lindane_ww, data_DDT_ww)  |> 
  left_join(esteem.overview::data_ROCCHMV_bio, by = "ESTUARY")

OC_pesticides_names <- c("Gamma-HCH", "sum_DDT")

OC_pesticides <- data.frame(PARAMETRE_LIBELLE = OC_pesticides_names,
                   NQE_ng_gww = c(0.29, 1282))
```

## Identify first and last periods

Identify the first and last 5 years of measurement
```{r}
# Compute minimal and maximal year of measurement by compounds and estuary
min_max_year <- data_OC_pesticides_ww |> 
  group_by(PARAMETRE_LIBELLE, ESTUARY) |> 
  summarise(min_year = min(YEAR),
            max_year = max(YEAR), .groups = "drop")

# Identify the first and last 5 years of measurement
data_OC_pesticides_ww <- data_OC_pesticides_ww |> 
  left_join(min_max_year, by = c("PARAMETRE_LIBELLE", "ESTUARY"))  |> 
  mutate(periode = case_when(
    YEAR <= min_year + 5 ~ "First 5 years",
    YEAR >= max_year - 5 ~ "Last 5 years",
    TRUE ~ NA
  ))  |> 
  mutate(last_trend = case_when(
    YEAR >= max_year - 15 ~ "Last 15 years",
    TRUE ~ NA
  ))
```

# Time series graphs

## Last 15 years Spearman's correlation coefficient + test
```{r}
res_OC_pesticides_ww_trends <- get_last_trend_info(data = data_OC_pesticides_ww, 
                                                   norm = OC_pesticides)

writexl::write_xlsx(res_OC_pesticides_ww_trends, "../inst/results/data_contam/OC_pesticides/recent_trends_OC_pesticides_ww.xlsx")
```

```{r}
ggplot_OC_pesticides_ww <- ggplot(data_OC_pesticides_ww) +
  aes(x = YEAR,
      y = log(RESULTAT),
      colour = ESTUARY) +
  geom_line() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom") +
  facet_grid(vars(PARAMETRE_LIBELLE), scales = "free_y") +
  labs(colour = "Estuary", y = "Contamination level in ng/g ww", x = NULL) +
# Add arrows for significant last trends
  geom_text(
    data = res_OC_pesticides_ww_trends,
    aes(x = last_year+1, y = log(last_resultat), label = symbol), size = rel(6),
    family = "DejaVu Sans",   # specify the font that contains arrows
    show.legend = FALSE)+
# Add environmental threshold when existing => UC QS(HH)
  geom_text(
    data = res_OC_pesticides_ww_trends,
    aes(x = last_year+1, y = log(NQE_ng_gww)*1.3,
        label = sprintf("%.2f", NQE_ng_gww)),
    color = "red", size = 2.5
  ) +
  geom_hline(
    data = res_OC_pesticides_ww_trends,
    aes(yintercept = log(NQE_ng_gww)),
    color = "red", linetype = "dashed"
  ) +
  theme_esteem()

ggplot_OC_pesticides_ww

ggsave(plot = ggplot_OC_pesticides_ww, "../inst/results/data_contam/OC_pesticides/ggplot_OC_pesticides_ww_time_series.jpg",
       width = 10, height = 8, units = "cm")
```

# First vs last period graphs
```{r}
data_OC_pesticides_ww_norm_relative <- data_OC_pesticides_ww |> 
  left_join(OC_pesticides) |> 
  mutate(RESULTAT = RESULTAT / NQE_ng_gww * 100) |> 
  drop_na(RESULTAT)

res_kruskal_OC_pesticides <- get_kruskal_periods_results(
  data_with_periode = data_OC_pesticides_ww_norm_relative, 
  chemicals = OC_pesticides_names)
```

```{r}
plot_OC_pesticides_periode <- data_OC_pesticides_ww_norm_relative |> 
  drop_na(periode) |> # keep only data from first and last 5-years periods
  ggplot() +
  aes(x = ESTUARY, y = RESULTAT, colour = periode) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.9)) +
  scale_color_manual(values = c("purple", "deeppink")) +
  facet_grid(rows = vars(PARAMETRE_LIBELLE), scales = "free_y") +
  labs(x = NULL, y = "Deviation from limit value \n (% ng/g ww)", colour = "Period") +
  theme(legend.position = "bottom") +
# Add long term trends if significant
  geom_text(data = res_kruskal_OC_pesticides, 
            inherit.aes = FALSE, 
            aes(x = ESTUARY, y = 0.9*y_pos, label = trend), size = rel(6),
    family = "DejaVu Sans",   # specify the font that contains arrows
    show.legend = FALSE) +
  theme_esteem()

plot_OC_pesticides_periode

ggsave(plot = plot_OC_pesticides_periode, "../inst/results/data_contam/OC_pesticides/ggplot_OC_pesticides_ww_periods.jpg", 
       width = 10, height = 8, units = "cm")
```

## General levels of contamination
```{r}
get_general_levels(data = data_OC_pesticides_ww, family = "OC_pesticides", suffix = "_ww")
```


# General evaluation
```{r}
get_general_evaluation(family = "OC_pesticides",
                       suffix = "_ww",
                       res_family_trends = res_OC_pesticides_ww_trends,
                       res_kruskal = res_kruskal_OC_pesticides)
```
