---
title: "data_contamination_PAH"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{data_contamination_PAH}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(esteem.overview)
library(tidyverse)
library(here)
library(writexl)
```

# Get raw data for Polycyclic Aromatic Hydrocarbon (PAH) => thresholds in ng/g dw 
```{r}
PAH <- data.frame(PARAMETRE_LIBELLE = c("Fluoranthene", "Benzo-pyr.", "Anthracene", 
                                        "Naphtalene", "Phenanthrene", "Pyrene",
                                        "Benzo-anthra.",  "Benzo-peryl."),
                   EAC_OSPAR_ng_gdw = c(110, 600, 290, 340, 1700, 100, 80, 110)
)

data_PAH <- esteem.overview::data_ROCCHMV_cleaned |> ungroup() |> 
  filter(support == "Bivalve") |> 
  select(-c(SUPPORT_NIVEAU_PRELEVEMENT, support, niveau)) |> 
  mutate(PARAMETRE_LIBELLE = case_when(
    PARAMETRE_LIBELLE == "Fluoranthène" ~ "Fluoranthene",
    PARAMETRE_LIBELLE == "Benzo(a)pyrène" ~ "Benzo-pyr.",
    PARAMETRE_LIBELLE == "Anthracène" ~ "Anthracene",
    PARAMETRE_LIBELLE == "Naphtalène" ~ "Naphtalene",
    PARAMETRE_LIBELLE == "Phénanthrène" ~ "Phenanthrene",
    PARAMETRE_LIBELLE == "Pyrène" ~ "Pyrene",
    PARAMETRE_LIBELLE == "Benzo(a)anthracène" ~ "Benzo-anthra.",
    PARAMETRE_LIBELLE == "Benzo(g,h,i)pérylène" ~ "Benzo-peryl.",
  )) |> 
  filter(PARAMETRE_LIBELLE %in% PAH$PARAMETRE_LIBELLE) |> 
  left_join(esteem.overview::data_ROCCHMV_bio, by = "ESTUARY")
```

## Get the same unit of measurment

Different units:

- 1994-2016: µg.kg-1				
- 2018-2024: µg.kg-1, p.h. (eq ng.g-1, p.h.)
```{r}
data_PAH |> 
  distinct(UNITE)

# data_PAH |> 
#   mutate(YEAR = as.factor(YEAR)) |> 
#   distinct(YEAR, UNITE, ESTUARY, PARAMETRE_LIBELLE) |>  View(title = "PAH units")

# # Check difference dw ww for a common YEAR x ESTUARY x chemical
# data_PAH |> 
#   filter(YEAR == 2020, ESTUARY == "Loire", PARAMETRE_LIBELLE == "Benzo(a)anthracène")
```

Transform all results in ng/g dw (same as µg/kg):
```{r}
data_PAH_dw <- data_PAH |>
  # transformation into dw concentrations
  mutate(RESULTAT = case_when(
    UNITE == "µg.kg-1, p.h." ~ RESULTAT / (`Matière sèche_%` / 100),
    UNITE == "µg.kg-1" ~ RESULTAT
  )) |>  
  # Change to unified unit
  mutate(UNITE = "ng/g dw") |> 
  # Get only one value per year
  group_by(ESTUARY, YEAR, PARAMETRE_LIBELLE) |>
  # Round result values
  mutate(RESULTAT = round(RESULTAT, digits = 3))
```

## Importance of the species matrix

- Gironde : only C. gigas
- Seine : only M. edulis
- Loire : 
  - 1994-2021: M. edulis
  - 2022-2024: M. edulis + M.galloprovincialis
  - 2018, 2021, 2024: Crassostrea gigas
```{r}
data_PAH_dw |> 
  distinct(ESTUARY, espece)

data_PAH_dw |> 
  filter(ESTUARY == "Loire") |> 
  distinct(ESTUARY, YEAR, espece)
```

No species effect
```{r}
data_PAH_dw |> 
  filter(ESTUARY == "Loire", YEAR > 2010) |>
  ggplot() +
  aes(x = YEAR, y = RESULTAT, colour = espece) +
  geom_line() +
  facet_wrap(vars(PARAMETRE_LIBELLE), scales = "free_y") +
  theme_esteem()
```

Simplify with two species groups
```{r}
data_PAH_dw <- data_PAH_dw |> 
    mutate(SPECIES = case_when(
    espece == "Crassostrea gigas (huître creuse)" ~ "oyster",
    espece %in% c("Mytilus edulis (moule commune)",
                  "Mytilus edulis + galloprovincialis (moule)") ~ "mussel"
  ))
```

## Identify first and last periods

Identify the first and last 5 years of measurement
```{r}
# Compute minimal and maximal year of measurement by compounds and estuary
min_max_year <- data_PAH_dw |> 
  group_by(PARAMETRE_LIBELLE, ESTUARY) |> 
  summarise(min_year = min(YEAR),
            max_year = max(YEAR), .groups = "drop")

# Identify the first and last 5 years of measurement
data_PAH_dw <- data_PAH_dw |> 
  left_join(min_max_year, by = c("PARAMETRE_LIBELLE", "ESTUARY"))  |> 
  mutate(periode = case_when(
    YEAR <= min_year + 5 ~ "First 5 years",
    YEAR >= max_year - 5 ~ "Last 5 years",
    TRUE ~ NA
  )) |> 
  mutate(last_trend = case_when(
    YEAR >= max_year - 15 ~ "Last 15 years",
    TRUE ~ NA
  ))
```

# Time series graphs

## Last 15 years Spearman's correlation coefficient + test
```{r}
res_PAH_dw_trends <- get_last_trend_info(data = data_PAH_dw, norm = PAH)

writexl::write_xlsx(res_PAH_dw_trends, "../inst/results/data_contam/PAH/recent_trends_PAH_ng_dw.xlsx")
```

```{r}
ggplot_PAH_dw <- ggplot(data_PAH_dw) +
  aes(x = YEAR,
      y = log(RESULTAT),
      colour = ESTUARY) +
  geom_line() +
  geom_smooth(se = FALSE) +
  facet_grid(rows = vars(
    fct_relevel(PARAMETRE_LIBELLE, PAH$PARAMETRE_LIBELLE)
  ), scales = "free_y") +
  theme(
    legend.position = "bottom",
    strip.text = element_text(size = 6.5) # facet title size
    ) +
  labs(colour = "Estuary", y = "Contamination level in log ng/g dw", x = NULL) +
# Add arrows for significant last trends
  geom_text(
    data = res_PAH_dw_trends,
    aes(x = last_year+1, y = log(last_resultat), label = symbol), size = rel(6),
    family = "DejaVu Sans",   # specify the font that contains arrows
    show.legend = FALSE) +
# Add environmental threshold when existing => EAC OSPAR 2023
  geom_text(
    data = res_PAH_dw_trends,  
    aes(x = last_year+1, y = max(log(last_resultat))*1.1, label = sprintf("%.2f", log(EAC_OSPAR_ng_gdw))),
    color = "red", size = 2.5
  ) +
  geom_hline(
    data = res_PAH_dw_trends,
    aes(yintercept = log(EAC_OSPAR_ng_gdw)),
    color = "red", linetype = "dashed"
  ) +
  theme_esteem()

ggplot_PAH_dw

ggsave(plot = ggplot_PAH_dw, "../inst/results/data_contam/PAH/ggplot_PAH_dw_time_series.jpg",
       width = 10, height = 15, units = "cm")
```

# First vs last period graphs

```{r}
data_PAH_dw_norm_relative <- data_PAH_dw |> 
  left_join(PAH) |> 
  mutate(RESULTAT = RESULTAT / EAC_OSPAR_ng_gdw * 100)

res_kruskal_PAH_dw <- get_kruskal_periods_results(
  data_with_periode = data_PAH_dw_norm_relative, 
  chemicals = PAH$PARAMETRE_LIBELLE
  )
```

```{r}
plot_PAH_dw_periode <- data_PAH_dw_norm_relative |> 
  drop_na(periode) |> # keep only data from first and last 5-years periods
  ggplot() +
  aes(x = ESTUARY, y = RESULTAT, colour = periode) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.9)) +
  scale_color_manual(values = c("purple", "deeppink")) +
  facet_grid(rows = vars(fct_relevel(PARAMETRE_LIBELLE, PAH$PARAMETRE_LIBELLE)), scales = "free_y") +
  labs(x = NULL, y = "Deviation from limit value (% ng/g dw)", colour = "Period") +
  theme(
    legend.position = "bottom",
    strip.text = element_text(size = 6.5) # facet title size
    ) +
# Add long term trends if significant
  geom_text(data = res_kruskal_PAH_dw, 
            inherit.aes = FALSE, 
            aes(x = ESTUARY, y = 0.9*y_pos, label = trend), size = rel(6),
    family = "DejaVu Sans",   # specify the font that contains arrows
    show.legend = FALSE) +
  theme_esteem()

plot_PAH_dw_periode

ggsave(plot = plot_PAH_dw_periode, "../inst/results/data_contam/PAH/ggplot_PAH_dw_periods.jpg", 
       width = 10, height = 15, units = "cm")
```

## General levels of contamination

```{r}
get_general_levels(data = data_PAH_dw, family = "PAH", suffix = "_dw")
```

# General evaluation

```{r}
get_general_evaluation(family = "PAH",
                       suffix = "_dw",
                       res_family_trends = res_PAH_dw_trends,
                       res_kruskal = res_kruskal_PAH_dw)
```
