---
title: "data_contamination_tributyltin"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{data_contamination_tributyltin}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
library(esteem.overview)
library(segmented)
library(tidyverse)
library(here)
```

# Get raw data for Polychlorinated biphenyls indicators (tributyltin) => thresholds in ng/g lw
```{r}
tributyltin <- data.frame(PARAMETRE_LIBELLE = "Tributylétain cation",
                   EAC_ng_gdw = 12)

data_tributyltin <- esteem.overview::data_ROCCHMV_cleaned |> ungroup() |> 
  filter(support == "Bivalve") |> 
  select(-c(SUPPORT_NIVEAU_PRELEVEMENT, support, niveau)) |> 
  filter(PARAMETRE_LIBELLE %in% tributyltin$PARAMETRE_LIBELLE) |> 
  left_join(esteem.overview::data_ROCCHMV_bio, by = "ESTUARY")

```

## Get the same unit of measurment

Different units:

- 2019-2020: µg/(kg MS)				
- 2021-2024: µg.kg-1, p.h.
```{r}
data_tributyltin |> 
  distinct(UNITE)

data_tributyltin |> 
  mutate(YEAR = as.factor(YEAR)) |> 
  distinct(YEAR, UNITE, ESTUARY, PARAMETRE_LIBELLE) |>  View(title = "tributyltin units")
```

Transform all results in ng/g ww (same as µg/kg):
```{r}
data_tributyltin_dw <- data_tributyltin |>
  # transformation into dw concentrations
  mutate(RESULTAT = case_when(
    UNITE == "µg.kg-1, p.h." ~ RESULTAT / (`Matière sèche_%` / 100),
    UNITE == "µg/(kg MS)" ~ RESULTAT ,
  )) |>  
  # Change to unified unit
  mutate(UNITE = "ng/g dw") |> 
  # Get only one value per year
  group_by(ESTUARY, YEAR, PARAMETRE_LIBELLE) |>
  mutate(RESULTAT = mean(RESULTAT)) |> 
  ungroup() |> 
  # Round result values
  mutate(RESULTAT = round(RESULTAT, digits = 3))

```

```{r}
data_tributyltin_dw |> 
  group_by(ESTUARY, PARAMETRE_LIBELLE) |> 
  summarise(mean = mean(RESULTAT))
```


## Importance of the species matrix

- Gironde : only C. gigas
- Seine : only M. edulis
- Loire : 
  - 2021: M. edulis
  - 2022-2024: M. edulis + M.galloprovincialis
```{r}
data_tributyltin_dw |> 
  distinct(ESTUARY, espece)

data_tributyltin_dw |> 
  filter(ESTUARY == "Loire") |> 
  distinct(ESTUARY, YEAR, espece)
```

```{r}
data_tributyltin_dw |> 
  filter(ESTUARY == "Loire") |>
  ggplot() +
  aes(x = YEAR, y = RESULTAT, colour = espece) +
  geom_line() +
  facet_wrap(vars(PARAMETRE_LIBELLE), scales = "free_y")
```

Simplify with two species groups
```{r}
data_tributyltin_dw <- data_tributyltin_dw |> 
    mutate(SPECIES = case_when(
    espece == "Crassostrea gigas (huître creuse)" ~ "oyster",
    espece %in% c("Mytilus edulis (moule commune)",
                  "Mytilus edulis + galloprovincialis (moule)") ~ "mussel"
  ))
```

# Time series graphs
```{r}
data_tributyltin_dw |>  
ggplot() +
  aes(x = YEAR, y = RESULTAT, color = ESTUARY) +
  geom_line() +
  facet_grid(vars(PARAMETRE_LIBELLE), scales = "free_y") +
  labs(x = NULL)
```

```{r}
fun_seg_model_tributyltin <- function(chemical, estuary){
  fun_seg_model(data = data_tributyltin_dw,
                folder = "tributyltin",
                chemical = chemical, 
                estuary = estuary)
}
  
tributyltin_names <- tributyltin$PARAMETRE_LIBELLE 
estuary_names <- c("Gironde", "Loire", "Seine")
params <- expand.grid(estuary_names, tributyltin_names)

res_tributyltin_dw_trends <- pmap_dfr(params, ~ fun_seg_model_tributyltin(estuary = ..1, chemical = ..2))

res_tributyltin_dw_trends <- res_tributyltin_dw_trends |> 
  left_join(tributyltin, by = "PARAMETRE_LIBELLE")
```

```{r}
ggplot_tributyltin_dw <- ggplot(data_tributyltin_dw) +
  aes(x = YEAR,
      y = RESULTAT,
      colour = ESTUARY) +
  geom_line() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom") +
  facet_grid(rows = vars(
    fct_relevel(PARAMETRE_LIBELLE, tributyltin_names)
  ), scales = "free_y") +
  labs(colour = "Estuary", y = "Contamination level in ng/g lw", x = NULL) +
# Add last break dates of segmented models and arrows for significant last trends
  geom_text(
    data = res_tributyltin_dw_trends,
    aes(x = max(last_year)+1, y = last_resultat, label = symbol), size = rel(6),
    family = "DejaVu Sans",   # specify the font that contains arrows
    show.legend = FALSE) +
# Add environmental threshold when existing
  geom_text(
    data = res_tributyltin_dw_trends,  
    aes(x = max(last_year)+1, y = EAC_ng_gdw +0.5, label = sprintf("%.2f", EAC_ng_gdw)),
    color = "red", size = 2.5
  ) +
  geom_hline(
    data = res_tributyltin_dw_trends,
    aes(yintercept = EAC_ng_gdw),
    color = "red", linetype = "dashed"
  )

ggplot_tributyltin_dw

ggsave(plot = ggplot_tributyltin_dw, "../inst/results/data_contam/tributyltin/ggplot_tributyltin_dw_time_series.jpg",
       width = 10, height = 10, units = "cm")
```

# First vs last period graphs
```{r}
data_tributyltin_dw_norm_relative <- data_tributyltin_dw |> 
  left_join(tributyltin) |> 
  mutate(RESULTAT = RESULTAT / EAC_ng_gdw * 100)
```

```{r}
plot_tributyltin_dw_periode <- data_tributyltin_dw_norm_relative |> 
  ggplot() +
  aes(x = ESTUARY, y = RESULTAT) +
  geom_boxplot(width = 0.6, colour = "deeppink") +
  facet_grid(rows = vars(PARAMETRE_LIBELLE), scales = "free_y") +
  labs(x = NULL, y = "Deviation from limit value (% ng/g dw)", colour = "Period") +
  theme(legend.position = "bottom")

plot_tributyltin_dw_periode

ggsave(plot = plot_tributyltin_dw_periode, "../inst/results/data_contam/tributyltin/ggplot_tributyltin_dw_periods.jpg", 
       width = 10, height = 10, units = "cm")
```

# General evaluation
```{r}
data_tributyltin_dw_norm_relative |> 
  group_by(ESTUARY, PARAMETRE_LIBELLE) |> 
  summarise(median = median(RESULTAT))
```

