---
title: "data_contamination_dioxine_like"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{data_contamination_dioxine_like}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(esteem.overview)
library(tidyverse)
library(here)
```

# Get raw data for Dioxine_like compounds => thresholds in ng/g TEQ
```{r}
dioxine_like_names <- tibble(PARAMETRE_LIBELLE = c(
  # DL-compounds: 12 DL-PCB
  "CB 77", "CB 81", "CB 105", "CB 114", "CB 118", "CB 123",
  "CB 126", "CB 156", "CB 157", "CB 167", "CB 169", "CB 189",
  # DL-compounds: 7 polychlorinated dibenzo-p-dioxins (PCDDs)
  "2,3,7,8-tetrachlorodibenzo-p-dioxine",
  "1,2,3,7,8-pentachlorodibenzo-p-dioxine",
  "1,2,3,4,7,8-hexachlorodibenzo-p-dioxine",
  "1,2,3,6,7,8-hexachlorodibenzo-p-dioxine",
  "1,2,3,7,8,9-hexachlorodibenzo-p-dioxine",
  "1,2,3,4,6,7,8-heptachlorodibenzo-p-dioxine",
  "octachlorodibenzo-p-dioxine", # OCDD 1,2,3,4,6,7,8,9-O8CDD
  # DL-compounds: 10 polychlorinated dibenzofurans (PCDFs)
  "2,3,7,8-tetrachlorodibenzofuran",
  "1,2,3,7,8-pentachlorodibenzofuran",
  "2,3,4,7,8-pentachlorodibenzofuran",
  "1,2,3,4,7,8-hexachlorodibenzofuran",
  "1,2,3,6,7,8-hexachlorodibenzofuran",
  "1,2,3,7,8,9-hexachlorodibenzofuran",
  "2,3,4,6,7,8-hexachlorodibenzofuran",
  "1,2,3,4,6,7,8-heptachlorodibenzofuran",
  "1,2,3,4,7,8,9-heptachlorodibenzofuran",
  "octachlorodibenzofuranne")) # OCDF 1,2,3,4,6,7,8,9-O8CDF

# Check if names in .xlsx are alike with the vector of names
dioxine_like <- dioxine_like_names |>
  left_join(esteem.overview::TEF_dioxine_like, by = c("PARAMETRE_LIBELLE" = "CHEMICAL"))


data_dioxine_like <- esteem.overview::data_ROCCHMV_cleaned |> ungroup() |> 
  filter(support == "Bivalve") |> 
  select(-c(SUPPORT_NIVEAU_PRELEVEMENT, support, niveau)) |> 
  filter(PARAMETRE_LIBELLE %in% dioxine_like$PARAMETRE_LIBELLE) |> 
  left_join(esteem.overview::data_ROCCHMV_bio, by = "ESTUARY") |> 
  left_join(TEF_dioxine_like, by = c("PARAMETRE_LIBELLE" = "CHEMICAL"))
```

## Get the same unit of measurment

Different units:

- 1992-2007: µg.kg-1 (eq ng.g-1)
- 2008-2016: PCDDs & PCDFs: ng.kg-1 ; PCBs: µg.kg-1
- 2017-2024: ng.g-1, p.h. 
```{r}
data_dioxine_like |> 
  distinct(UNITE)

# data_dioxine_like |> 
#   mutate(YEAR = as.factor(YEAR)) |> 
#   distinct(YEAR, UNITE, ESTUARY, PARAMETRE_LIBELLE) |>  View(title = "dioxine_like units")
```

Transform all results in ng/g dw (same as µg/kg):
```{r}
data_dioxine_like_dw <- data_dioxine_like |>
  # transformation into dw concentrations
  mutate(RESULTAT = case_when(
    UNITE == "ng.g-1, p.h." ~ RESULTAT / (`Matière sèche_%` / 100),
    UNITE == "µg.kg-1" ~ RESULTAT,
    UNITE == "ng.kg-1" ~ RESULTAT / 1000
  )) |>  
  # Change to unified unit
  mutate(UNITE = "ng/g dw") |> 
  # Get only one value per year
  group_by(ESTUARY, YEAR, PARAMETRE_LIBELLE) |>
  mutate(RESULTAT = mean(RESULTAT)) |> 
  ungroup() |> 
  # Round result values
  mutate(RESULTAT = round(RESULTAT, digits = 7))

# # Checking
# data_dioxine_like |> filter(YEAR %in% c(2015, 2023), PARAMETRE_LIBELLE == "octachlorodibenzofuranne")
# data_dioxine_like_dw |> filter(YEAR %in% c(2015, 2023), PARAMETRE_LIBELLE == "octachlorodibenzofuranne")
```

Transform all results in ng/g TEQ:
```{r}
data_dioxine_like_TEQ <- data_dioxine_like_dw |> 
  # transformation into lw concentrations
  mutate(RESULTAT = RESULTAT * `TEF OMS 2005`) |>  
  # Change to unified unit
  mutate(UNITE = "ng/g TEQ") |> 
  # Round result values
  mutate(RESULTAT = round(RESULTAT, digits = 7))

# # Checking => should be the same (TEF=1)
# data_dioxine_like_dw |> filter(YEAR %in% c(2015, 2023), PARAMETRE_LIBELLE == "1,2,3,7,8-pentachlorodibenzo-p-dioxine")
# data_dioxine_like_TEQ |> filter(YEAR %in% c(2015, 2023), PARAMETRE_LIBELLE == "1,2,3,7,8-pentachlorodibenzo-p-dioxine")

# # Checking => should be 10% (TEF=0.1)
# data_dioxine_like_dw |> filter(YEAR %in% c(2015, 2023), PARAMETRE_LIBELLE == "1,2,3,4,7,8-hexachlorodibenzo-p-dioxine")
# data_dioxine_like_TEQ |> filter(YEAR %in% c(2015, 2023), PARAMETRE_LIBELLE == "1,2,3,4,7,8-hexachlorodibenzo-p-dioxine")
```

## Get the monitoring periods
```{r}
data_dioxine_like_TEQ |> 
  group_by(ESTUARY, PARAMETRE_LIBELLE) |> 
  summarise(min_year = min(YEAR),
            max_year = max(YEAR))
``` 

## Compute the sum
```{r}
data_sum_dioxine_like_TEQ <- data_dioxine_like_TEQ |> 
  filter(YEAR >= 2008) |> 
  group_by(ESTUARY, YEAR, espece) |> 
  summarise(
    RESULTAT = sum(RESULTAT), 
    .groups = "drop") |> 
  mutate(PARAMETRE_LIBELLE = "sum_dioxine_like")

dioxine_like_norm <-  data.frame(PARAMETRE_LIBELLE = "sum_dioxine_like",
                                 QS_HH_EC_TEQ = 0.0065)
```


## Importance of the species matrix

- Gironde : only C. gigas
- Seine : only M. edulis
- Loire : 
  - 1992-2021: M. edulis
  - 2022-2024: M. edulis + M.galloprovincialis
  - 2018, 2021, 2024: Crassostrea gigas
```{r}
data_sum_dioxine_like_TEQ |> 
  distinct(ESTUARY, espece)

data_sum_dioxine_like_TEQ |> 
  filter(ESTUARY == "Loire") |> 
  distinct(ESTUARY, YEAR, espece)
```

Don't know what I can conclude from that:
```{r}
data_sum_dioxine_like_TEQ |> 
  filter(ESTUARY == "Loire", YEAR > 2010) |>
  ggplot() +
  aes(x = YEAR, y = RESULTAT, colour = espece) +
  geom_line() +
  facet_wrap(vars(PARAMETRE_LIBELLE), scales = "free_y")
```

Simplify with two species groups
```{r}
data_sum_dioxine_like_TEQ <- data_sum_dioxine_like_TEQ |> 
    mutate(SPECIES = case_when(
    espece == "Crassostrea gigas (huître creuse)" ~ "oyster",
    espece %in% c("Mytilus edulis (moule commune)",
                  "Mytilus edulis + galloprovincialis (moule)") ~ "mussel"
  ))
```

# Time series graphs

## Spearman's correlation coefficient + test
```{r}
# Compute for all time serie
data_sum_dioxine_like_TEQ <- data_sum_dioxine_like_TEQ |> 
  mutate(last_trend = "Last 15 years")

res_sum_dioxine_like_lw_trends <- get_last_trend_info(data = data_sum_dioxine_like_TEQ, norm = dioxine_like_norm)

writexl::write_xlsx(res_sum_dioxine_like_lw_trends, "../inst/results/data_contam/dioxine_like/recent_trends_dioxine_like_TEQ.xlsx")
```


```{r}
ggplot_dioxine_like_TEQ <- ggplot(data_sum_dioxine_like_TEQ) +
  aes(x = YEAR,
      y = RESULTAT,
      colour = ESTUARY) +
  geom_line() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom") +
  facet_grid(vars(PARAMETRE_LIBELLE)) +
  labs(colour = "Estuary", y = "Contamination level \n in ng/g TEQ", x = NULL) +
# Add arrows for significant last trends
  geom_text(
    data = res_sum_dioxine_like_lw_trends,
    aes(x = last_year+1, y = last_resultat, label = symbol), size = rel(6),
    family = "DejaVu Sans",   # specify the font that contains arrows
    show.legend = FALSE) +
# Add environmental threshold when existing => UC QS(HH)
  geom_text(
    data = res_sum_dioxine_like_lw_trends,  
    aes(x = last_year+1, y = 0.0065*1.1, label = "0.0065"),
    color = "red", size = 2.5
  ) +
  geom_hline(
    data = res_sum_dioxine_like_lw_trends,
    aes(yintercept = 0.0065),
    color = "red", linetype = "dashed"
  )

ggplot_dioxine_like_TEQ

ggsave(plot = ggplot_dioxine_like_TEQ, "../inst/results/data_contam/dioxine_like/ggplot_dioxine_like_TEQ_time_series.jpg",
       width = 10, height = 7, units = "cm")
```

# First vs last period graphs
```{r}
data_sum_dioxine_like_TEQ_norm_relative <- data_sum_dioxine_like_TEQ |> 
  mutate(QS_HH = 0.0065) |> 
  mutate(RESULTAT = RESULTAT / QS_HH * 100)
```

```{r}
plot_dioxine_like_periode <- data_sum_dioxine_like_TEQ_norm_relative |> 
  ggplot() +
  aes(x = ESTUARY, y = RESULTAT, colour = periode) +
  geom_boxplot(width = 0.6, colour = "deeppink") +
  facet_grid(rows = vars(PARAMETRE_LIBELLE), scales = "free_y") +
  labs(x = NULL, y = "Deviation from limit value \n (% ng/g TEQ)", colour = "Period") +
  theme(legend.position = "bottom")

plot_dioxine_like_periode

ggsave(plot = plot_dioxine_like_periode, "../inst/results/data_contam/dioxine_like/ggplot_dioxine_like_TEQ_periods.jpg", 
       width = 10, height = 7, units = "cm")
```

## General levels of contamination
```{r}
data_dioxine_like_TEQ_levels <- data_sum_dioxine_like_TEQ |> 
  group_by(ESTUARY, PARAMETRE_LIBELLE) |> 
  summarise(median_ng_gTEQ = median(RESULTAT), .groups = "drop") |> 
  arrange(PARAMETRE_LIBELLE)

data_dioxine_like_TEQ_norm_relative <- data_sum_dioxine_like_TEQ_norm_relative |> 
  group_by(ESTUARY, PARAMETRE_LIBELLE) |> 
  summarise(median_percent = median(RESULTAT), .groups = "drop") |> 
  arrange(PARAMETRE_LIBELLE) |>  
  select(ESTUARY, PARAMETRE_LIBELLE, median_percent)

data_dioxine_like_general_levels <- full_join(data_dioxine_like_TEQ_levels, data_dioxine_like_TEQ_norm_relative)

writexl::write_xlsx(x = data_dioxine_like_general_levels, path = "../inst/results/data_contam/dioxine_like/summarised_levels_data_dioxine_like_TEQ.xlsx")
```


# General evaluation

```{r}
# trends
res_sum_dioxine_like_lw_trends

# color dot from percentage
# blue: [0-50%[, green: [50-100%[, yellow:[100-150%[, orange: [150-300%[, red: [300-500%[, black: > 500%
data_sum_dioxine_like_TEQ_norm_relative |>
  group_by(ESTUARY, PARAMETRE_LIBELLE) |>
  summarise(median = median(RESULTAT))
```
