---
title: "data_contamination_PCBi"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{data_contamination_PCBi}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
library(esteem.overview)
library(segmented)
library(tidyverse)
library(here)
```

# Get raw data for Polychlorinated biphenyls indicators (PCBi) => thresholds in ng/g lw
```{r}
PCBi <- data.frame(PARAMETRE_LIBELLE = c("CB 28", "CB 52", "CB 101", "CB 118", "CB 138", "CB 153", "CB 180"),
                   EAC_OSPAR_ng_g_lw = c(67, 108, 121, 25, 317, 1585, 469))

data_PCBi <- esteem.overview::data_ROCCHMV_cleaned |> ungroup() |> 
  filter(support == "Bivalve") |> 
  select(-c(SUPPORT_NIVEAU_PRELEVEMENT, support, niveau)) |> 
  filter(PARAMETRE_LIBELLE %in% PCBi$PARAMETRE_LIBELLE) |> 
  left_join(esteem.overview::data_ROCCHMV_bio, by = "ESTUARY")
```

## Get the same unit of measurment

Different units:

- 1993-2016: µg.kg-1				
- 2017-2024: ng.g-1, p.h. (eq µg.kg-1, p.h.)
```{r}
data_PCBi |> 
  distinct(UNITE)

# data_PCBi |> 
#   mutate(YEAR = as.factor(YEAR)) |> 
#   distinct(YEAR, UNITE, ESTUARY, PARAMETRE_LIBELLE) |>  View(title = "PCBi units")
```

Transform all results in ng/g dw (same as µg/kg):
```{r}
data_PCBi_dw <- data_PCBi |>
  # transformation into dw concentrations
  mutate(RESULTAT = case_when(
    UNITE == "ng.g-1, p.h." ~ RESULTAT / (`Matière sèche_%` / 100),
    UNITE == "µg.kg-1" ~ RESULTAT
  )) |>  
  # Change to unified unit
  mutate(UNITE = "ng/g dw") |> 
  # Get only one value per year
  group_by(ESTUARY, YEAR, PARAMETRE_LIBELLE) |>
  mutate(RESULTAT = mean(RESULTAT)) |> 
  ungroup() |> 
  # Round result values
  mutate(RESULTAT = round(RESULTAT, digits = 3))

# # Checking
# data_PCBi |> filter(YEAR %in% c(2015, 2023), PARAMETRE_LIBELLE == "CB 153")
# data_PCBi_dw |> filter(YEAR %in% c(2015, 2023), PARAMETRE_LIBELLE == "CB 153")
```

Transform all results in ng/g lw:
```{r}
data_PCBi_lw <- data_PCBi_dw |> 
  # transformation into lw concentrations
  mutate(RESULTAT = case_when(
    UNITE == "ng/g dw" ~ RESULTAT / (`Lipides totaux_%` / 100),
    TRUE ~ RESULTAT
  )) |>  
  # Change to unified unit
  mutate(UNITE = "ng/g lw") |> 
  # Round result values
  mutate(RESULTAT = round(RESULTAT, digits = 3))

# # Checking
# data_PCBi_dw |> filter(YEAR %in% c(2015, 2023), PARAMETRE_LIBELLE == "CB 153")
# data_PCBi_lw |> filter(YEAR %in% c(2015, 2023), PARAMETRE_LIBELLE == "CB 153")
```

## Importance of the species matrix

- Gironde : only C. gigas
- Seine : only M. edulis
- Loire : 
  - 1992-2021: M. edulis
  - 2022-2023: M. edulis + M.galloprovincialis
  - 2018, 2021, 2024: Crassostrea gigas
```{r}
data_PCBi_lw |> 
  distinct(ESTUARY, espece)

data_PCBi_lw |> 
  filter(ESTUARY == "Loire") |> 
  distinct(ESTUARY, YEAR, espece)
```

```{r}
data_PCBi_lw |> 
  filter(ESTUARY == "Loire", YEAR > 2010) |>
  ggplot() +
  aes(x = YEAR, y = RESULTAT, colour = espece) +
  geom_line() +
  facet_wrap(vars(PARAMETRE_LIBELLE), scales = "free_y")
```

Simplify with two species groups
```{r}
data_PCBi_lw <- data_PCBi_lw |> 
    mutate(SPECIES = case_when(
    espece == "Crassostrea gigas (huître creuse)" ~ "oyster",
    espece %in% c("Mytilus edulis (moule commune)",
                  "Mytilus edulis + galloprovincialis (moule)") ~ "mussel"
  ))
```

## Identify first and last periods

Identify the first and last 5 years of measurement
```{r}
# Compute minimal and maximal year of measurement by compounds and estuary
min_max_year <- data_PCBi_lw |> 
  group_by(PARAMETRE_LIBELLE, ESTUARY) |> 
  summarise(min_year = min(YEAR),
            max_year = max(YEAR), .groups = "drop")

# Identify the first and last 5 years of measurement
data_PCBi_lw <- data_PCBi_lw |> 
  left_join(min_max_year, by = c("PARAMETRE_LIBELLE", "ESTUARY"))  |> 
  mutate(periode = case_when(
    YEAR <= min_year + 5 ~ "First 5 years",
    YEAR >= max_year - 5 ~ "Last 5 years",
    TRUE ~ NA
  ))
```

# Time series graphs
```{r}
data_PCBi_lw |>  
ggplot() +
  aes(x = YEAR, y = RESULTAT, color = ESTUARY) +
  geom_line() +
  facet_grid(vars(PARAMETRE_LIBELLE), scales = "free_y") +
  labs(x = NULL)
```

```{r}
fun_seg_model_PCBi <- function(chemical, estuary){
  fun_seg_model(data = data_PCBi_lw,
                folder = "PCBi",
                chemical = chemical, 
                estuary = estuary)
}
  
PCBi_names <- PCBi$PARAMETRE_LIBELLE 
estuary_names <- c("Gironde", "Loire", "Seine")
params <- expand.grid(estuary_names, PCBi_names)

res_PCBi_lw_trends <- pmap_dfr(params, ~ fun_seg_model_PCBi(estuary = ..1, chemical = ..2))

res_PCBi_lw_trends <- res_PCBi_lw_trends |> 
  left_join(PCBi, by = "PARAMETRE_LIBELLE")
```

```{r}
ggplot_PCBi_lw <- ggplot(data_PCBi_lw) +
  aes(x = YEAR,
      y = log(RESULTAT),
      colour = ESTUARY) +
  geom_line() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom") +
  facet_grid(rows = vars(
    fct_relevel(PARAMETRE_LIBELLE, PCBi_names)
  ), scales = "free_y") +
  labs(colour = "Estuary", y = "Contamination level in ng/g lw", x = NULL) +
# Add last break dates of segmented models and arrows for significant last trends
  geom_vline(data = res_PCBi_lw_trends, aes(xintercept = break_dates, colour = ESTUARY)) +
  geom_text(
    data = res_PCBi_lw_trends,
    aes(x = last_year+1, y = log(last_resultat), label = symbol), size = rel(6),
    family = "DejaVu Sans",   # specify the font that contains arrows
    show.legend = FALSE) +
# Add environmental threshold when existing => EAC OSPAR 2023
  geom_text(
    data = res_PCBi_lw_trends,  
    aes(x = last_year+1, y = max(log(last_resultat))*1.1, label = sprintf("%.2f", log(EAC_OSPAR_ng_g_lw))),
    color = "red", size = 2.5
  ) +
  geom_hline(
    data = res_PCBi_lw_trends,
    aes(yintercept = log(EAC_OSPAR_ng_g_lw)),
    color = "red", linetype = "dashed"
  )

ggplot_PCBi_lw

ggsave(plot = ggplot_PCBi_lw, "../inst/results/data_contam/PCBi/ggplot_PCBi_lw_time_series.jpg",
       width = 10, height = 15, units = "cm")
```

# First vs last period graphs
```{r}
data_PCBi_lw_norm_relative <- data_PCBi_lw |> 
  left_join(PCBi) |> 
  mutate(RESULTAT = RESULTAT / EAC_OSPAR_ng_g_lw * 100)

res_kruskal_PCBi_lw <- get_kruskal_periods_results(data_with_periode = data_PCBi_lw_norm_relative, 
                             chemicals = PCBi$PARAMETRE_LIBELLE)
```

```{r}
plot_PCBi_lw_periode <- data_PCBi_lw_norm_relative |> 
  drop_na(periode) |> # keep only data from first and last 5-years periods
  ggplot() +
  aes(x = ESTUARY, y = RESULTAT, colour = periode) +
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.9)) +
  scale_color_manual(values = c("purple", "deeppink")) +
  facet_grid(rows = vars(fct_relevel(PARAMETRE_LIBELLE, PCBi$PARAMETRE_LIBELLE)), scales = "free_y") +
  labs(x = NULL, y = "Deviation from limit value (% ng/g lw)", colour = "Period") +
  theme(legend.position = "bottom") +
# Add long term trends if significant
  geom_text(data = res_kruskal_PCBi_lw, 
            inherit.aes = FALSE, 
            aes(x = ESTUARY, y = 0.9*y_pos, label = trend), size = rel(6),
    family = "DejaVu Sans",   # specify the font that contains arrows
    show.legend = FALSE)

plot_PCBi_lw_periode

ggsave(plot = plot_PCBi_lw_periode, "../inst/results/data_contam/PCBi/ggplot_PCBi_lw_periods.jpg", 
       width = 10, height = 15, units = "cm")
```

## General levels of contamination
```{r}
data_PCBi_lw_levels <- data_PCBi_lw |> 
  filter(periode == "Last 5 years") |> 
  group_by(ESTUARY, PARAMETRE_LIBELLE) |> 
  summarise(median_ng_glw = median(RESULTAT), .groups = "drop") |> 
  arrange(PARAMETRE_LIBELLE)

data_PCBi_lw_norm_relative <- res_kruskal_PCBi_lw |> 
  select(ESTUARY, PARAMETRE_LIBELLE, median_1, median_2)

data_PCBi_general_levels <- full_join(data_PCBi_lw_levels, data_PCBi_lw_norm_relative)

writexl::write_xlsx(x = data_PCBi_general_levels, path = "../inst/results/data_contam/PCBi/summarised_levels_data_PCBi_lw.xlsx")
```


# General evaluation
```{r}
get_general_evaluation(family = "PCBi",
                       suffix = "_lw",
                       res_family_trends = res_PCBi_lw_trends,
                       res_kruskal = res_kruskal_PCBi_lw)
```

