---
title: "data_benthos"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{data_benthos}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
knitr::opts_chunk$set(echo = FALSE)
library(esteem.overview)
library(tidyverse, quietly = TRUE)
library(writexl)
# Maps
library(sf)
# Get taxonomy api
library(jsonlite) # install.packages("jsonlite", repos="http://cran.r-project.org")
library(httr) # install.packages("httr")
# Get benthos indicator AMBI
library(benthos)

`%!in%` = Negate(`%in%`)
```

# REBENT Indicators

```{r}
data_REBENT <- esteem.overview::data_benthos_REBENT_cleaned
```


## Taxonomic composition

### Phylum level

ni = Number of individuals for each phylum per estuary, tidal zone and year
```{r}
nb_ind_sp <- data_REBENT |> 
  group_by(ESTUARY, tidal, YEAR, phylum) |> 
  summarise(n_ind_sp = sum(RESULTAT), .groups = "drop")
```
N = Total number of individuals per estuary, tidal zone and year
```{r}
nb_ind_tot <- data_REBENT |> 
  group_by(ESTUARY, tidal, YEAR) |> 
  summarise(n_ind_tot = sum(RESULTAT), .groups = "drop")
```
pi = Percentage abundance for each phylum per estuary, tidal zone and year
```{r}
abundance_prop <- nb_ind_sp |> 
  left_join(nb_ind_tot, by = c("ESTUARY", "tidal", "YEAR")) |> 
  group_by(ESTUARY, tidal, YEAR) |> 
  mutate(abundance_prop = n_ind_sp / n_ind_tot * 100) |> 
  ungroup() |> 
  drop_na()
```

```{r}
ggplot_REBENT_abundance_taxonomy <- abundance_prop |> 
  filter(phylum %in% c("Annelida", "Arthropoda", "Mollusca")) |> 
ggplot() +
  aes(x = YEAR, y = abundance_prop, color = phylum, fill = phylum) +
  geom_col() +
  facet_grid(rows = vars(ESTUARY), cols = vars(tidal)) +
  labs(title = "REBENT macrobenthic fauna: Taxonomic repartition", 
       x = NULL, 
       y = "Abundance (%)",
       color = "Phylum", fill = "Phylum")

ggplot_REBENT_abundance_taxonomy

ggsave(plot = ggplot_REBENT_abundance_taxonomy,
       "../inst/results/data_benthos/indicators/ggplot_REBENT_abundance_taxonomy.jpg",
       width = 15, height = 6, units = "cm")
```

### Species level
```{r}
nb_ind_tot <- data_REBENT |> 
  group_by(ESTUARY, tidal) |> 
  summarise(n_ind_tot = sum(RESULTAT), .groups = "drop")

nb_ind_sp <- data_REBENT |> 
  group_by(ESTUARY, tidal, phylum, class, order, species) |> 
  summarise(n_ind_sp = sum(RESULTAT), .groups = "drop")

abundance_prop <- nb_ind_sp |> 
  left_join(nb_ind_tot, by = c("ESTUARY", "tidal")) |> 
  group_by(ESTUARY, tidal, phylum, class, order) |> 
  mutate(abundance_prop = n_ind_sp / n_ind_tot * 100) |> 
  ungroup() |> 
  drop_na()
```


```{r}
abundance_prop_species <- abundance_prop |> 
  filter(abundance_prop > 10) |> 
  arrange(ESTUARY, tidal, phylum, class, order, desc(abundance_prop)) |> 
  select(-c(n_ind_sp, n_ind_tot))
abundance_prop_species

write_xlsx(abundance_prop_species, "../inst/results/data_benthos/abundance_prop_species.xlsx")
```

```{r}
abundance_prop_species_5 <- abundance_prop |> 
  filter(abundance_prop > 5) |> 
  arrange(ESTUARY, tidal, phylum, class, order, desc(abundance_prop)) |> 
  select(-c(n_ind_sp, n_ind_tot))
abundance_prop_species_5

ggplot_REBENT_abundance_species <- abundance_prop_species_5 |> 
  filter(phylum %in% c("Annelida", "Arthropoda", "Mollusca")) |>
  mutate(species_grouped = paste(phylum, species, sep=": ")) |> 
  ggplot() +
  aes(x = phylum, y = abundance_prop, fill = species_grouped) +
  geom_col() +
  facet_grid(rows = vars(ESTUARY), cols = vars(tidal)) +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(ncol = 2)) +
  scale_fill_manual(values = c("pink", "forestgreen", "lightblue", "gold1", "red", "blue", "purple", "green", "orange", "yellow", "gold4", "blue4", "coral", "grey", "royalblue", "pink3","green2", "orange3", "black"),
                    guide = guide_legend(title = "Species (by Phylum)")) +
  labs(title = "REBENT macrobenthic fauna: Species composition", 
       x = NULL, 
       y = "Abundance (%)",
       fill = "Species")

ggplot_REBENT_abundance_species

ggsave(plot = ggplot_REBENT_abundance_species, "../inst/results/data_benthos/indicators/ggplot_REBENT_abundance_species.jpg",height = 10, width = 8)
```


## Species richness (S)

```{r}
species_richness <- data_REBENT |> 
  group_by(ESTUARY, YEAR, tidal, species) |> 
  summarise(count = sum(RESULTAT)) |> 
  summarise(sp_rich = species_richness(taxon = species, count = count), .groups = "drop")

ggplot_REBENT_sp_richness <- ggplot(species_richness) +
  aes(x = YEAR, y = sp_rich, color = ESTUARY,
      linetype = tidal, shape = tidal) +
  geom_point() +
  geom_line() +
  labs(title = "REBENT macrobenthic fauna - Species richness", 
       x = NULL, y = "Species richness",
       color = "Estuary")
ggplot_REBENT_sp_richness
ggsave(plot = ggplot_REBENT_sp_richness, filename = "../inst/results/data_benthos/indicators/ggplot_REBENT_species_richness.jpg",
       width = 15, height = 6, units = "cm")
```


## Shannon index (H')

ni = Number of individuals for each species per estuary, tidal zone and year
```{r}
nb_ind_sp <- data_REBENT |> 
  group_by(ESTUARY, tidal, YEAR, species) |> 
  summarise(n_ind_sp = sum(RESULTAT), .groups = "drop")
```

N = Total number of individuals per estuary, tidal zone and year
```{r}
nb_ind_tot <- data_REBENT |> 
  group_by(ESTUARY, tidal, YEAR) |> 
  summarise(n_ind_tot = sum(RESULTAT), .groups = "drop")
```

pi = Percentage abundance for each species per estuary, tidal zone and year
```{r}
abundance_prop <- nb_ind_sp |> 
  left_join(nb_ind_tot, by = c("ESTUARY", "tidal", "YEAR")) |> 
  group_by(ESTUARY, tidal, YEAR) |> 
  mutate(abundance_prop = n_ind_sp / n_ind_tot) |> 
  ungroup()
```

H' = Shannon-index
```{r}
shannon_index <- abundance_prop |> 
  group_by(ESTUARY, tidal, YEAR) |> 
  summarise(shannon = - sum(abundance_prop * log2(abundance_prop)), .groups = "drop")
```

```{r}
ggplot_shannon_index <- ggplot(shannon_index) +
  aes(x = YEAR, y = shannon, color = ESTUARY,
      linetype = tidal, shape = tidal) +
  geom_point() +
  geom_line() +
  labs(title = "REBENT macrobenthic fauna - Shannon index", 
       x = NULL, y = "Shannon index",
       color = "Estuary")
ggplot_shannon_index
ggsave(plot = ggplot_shannon_index, filename = "../inst/results/data_benthos/indicators/ggplot_REBENT_shannon_index.jpg",
       width = 15, height = 6, units = "cm")
```


## Abundance/Density

### Number of individual per m2 at a given LIEU_MNEMONIQUE & DATE

One line per species: nb of individuals
```{r}
data_benthos_ind_species_m2 <- data_REBENT |> 
  # nb of sub-sample per LIEU_MNEMONIQUE & DATE & sampler_type
  group_by(ESTUARY, LIEU_MNEMONIQUE, DATE, sampler_type, tidal) |> 
  mutate(nb_samples_lieu_mnemonique_date = unique(mnemonique) |>  length()) |> 
  # summarise nb of ind per species at a given LIEU_MNEMONIQUE & DATE & sampler_type
  group_by(ESTUARY, LIEU_MNEMONIQUE, DATE, sampler_type, sampler_surface,
           tidal, nb_samples_lieu_mnemonique_date, species) |> 
  summarise(nb_ind_species_lieu_mnemonique_date = sum(RESULTAT)) |> 
  # compute nb of in per species for the total surface of LIEU_MNEMONIQUE 
  mutate(nb_ind_species_per_m2 = nb_ind_species_lieu_mnemonique_date / (sampler_surface * nb_samples_lieu_mnemonique_date)) |> 
  ungroup()

data_benthos_ind_species_m2
```

Yearly mean of individuals per surface
```{r}
data_benthos_ind_m2 <- data_benthos_ind_species_m2 |> 
  mutate(YEAR = year(DATE)) |> 
  group_by(ESTUARY, YEAR, DATE, LIEU_MNEMONIQUE, tidal) |> 
  summarise(nb_ind_per_m2 = sum(nb_ind_species_per_m2)) |> 
  group_by(ESTUARY, YEAR, tidal) |> 
  summarise(mean_abundance_m2 = mean(nb_ind_per_m2)) |> 
  ungroup()
```


```{r}
ggplot_REBENT_density <- ggplot(data_benthos_ind_m2) +
  aes(x = YEAR, y = mean_abundance_m2, color = ESTUARY,
      linetype = tidal, shape = tidal) +
  geom_line(position = "dodge") +
  geom_point() +
  labs(title = "REBENT macrobenthic fauna: Density", 
       x = NULL, 
       y = "Density (mean nb ind/m²)",
       color = "Estuary")

ggsave(plot = ggplot_REBENT_density,
       "../inst/results/data_benthos/indicators/ggplot_REBENT_density.jpg",
       width = 17, height = 10, units = "cm")

ggplot_REBENT_density
```

### Species richness per square metre

```{r}
# Number of species per sample
data_n_species_by_sample <- data_benthos_ind_species_m2 |> 
  group_by(ESTUARY, LIEU_MNEMONIQUE, DATE, tidal) |> 
  summarise(n_species = n(), .groups = "drop")

# Sample surface
data_sample_surface <- data_benthos_ind_species_m2 |> 
  distinct(ESTUARY, LIEU_MNEMONIQUE, DATE, sampler_type, sampler_surface, tidal, nb_samples_lieu_mnemonique_date) |> 
  mutate(surface_sample = sampler_surface * nb_samples_lieu_mnemonique_date)

# Compute species richness by m²
data_species_richness <- left_join(data_n_species_by_sample, data_sample_surface) |> 
  mutate(n_species_per_m2 = n_species / surface_sample)

# Summarise per estuary and year
data_species_richness_year <- data_species_richness |> 
  mutate(YEAR = year(DATE)) |> 
  group_by(ESTUARY, YEAR, tidal) |> 
  summarise(mean_year_n_species_per_m2 = mean(n_species_per_m2), 
            sum_sample_surface = sum(surface_sample), .groups = "drop")
```

Impact of total surface sampled variability on estimated species richness:
```{r}
## Total surface sampled variability
  ggplot_surface_variability <- ggplot(data = data_species_richness_year) +
    aes(x = YEAR, y = sum_sample_surface, color = ESTUARY,
        linetype = tidal, shape = tidal) +
    geom_point() + geom_line() +
    labs(x = NULL, y = "Total surface sampled (m²)", 
         title = "Rebent macrobenthic fauna: Total surface sampled variability")
  ggplot_surface_variability
  ggsave(plot = ggplot_surface_variability, filename = "../inst/results/data_benthos/indicators/ggplot_REBENT_surface_variability.jpg",
         width = 15, height = 6, units = "cm")

  # ## Correlation between species richness and total surface sampled
  # model_lm <- lm(data = data_species_richness_year, 
  #    formula = mean_year_n_species_per_m2 ~ sum_sample_surface)
  # summary(model_lm) 
  
  ggplot_surface_richness_correlation <- ggplot(data = data_species_richness_year) +
    aes(x = sum_sample_surface, y = mean_year_n_species_per_m2, 
        color = ESTUARY, linetype = tidal, shape = tidal) +
    geom_point() +
    labs(x = "Total surface sampled", y = "Species richness",
         title = "Correlation between yearly total surface sampled and species richness")
  ggplot_surface_richness_correlation
  ggsave(plot = ggplot_surface_richness_correlation, filename = "../inst/results/data_benthos/indicators/ggplot_REBENT_surface_richness_correlation.jpg",
           width = 15, height = 6, units = "cm")
```

Plot of species richness per m² per estuary and year:
```{r}
ggplot_species_richness_year <- ggplot(data_species_richness_year) +
  aes(x = YEAR, y = mean_year_n_species_per_m2, color = ESTUARY,
      linetype = tidal, shape = tidal) +
  geom_point() +
  geom_line() +
  labs(title = "REBENT macrobenthic fauna - Species richness per square metre", 
       x = NULL, y = "Mean number of species per m²",
       color = "Estuary")
ggplot_species_richness_year
ggsave(plot = ggplot_species_richness_year, filename = "../inst/results/data_benthos/indicators/ggplot_REBENT_species_richness_squarem.jpg",
       width = 15, height = 6, units = "cm")
```

## AZTI Marine Biotic Index (AMBI)

Ecological groups of sensitivity to anthropic disturbance
```{r}
# List of ESTEEM species
list_species_REBENT <- data_REBENT |> 
  distinct(ESTUARY, YEAR, AphiaID, phylum, class, order, species) |> 
  mutate(species = case_when(
    species == "Nemertes" ~ "Nemertea",
    TRUE ~ species))

#  for benthos ecological groups
List_Group_Taxon_AMBI <- left_join(
  data.frame(readRDS(system.file("extdata", "azti.rds", package = "benthos"))), # DB Azti
  as.data.frame(get_ambi(which = "NL")))  # Dutch DB


# Joining
grp_taxon_ambi <- list_species_REBENT |> left_join(List_Group_Taxon_AMBI, by = c("species" = "TAXON"))
  

# Get percentage of species with AMBI ecological group
percent_without_ambi_grp_taxon <- grp_taxon_ambi |> 
  group_by(ESTUARY, YEAR) |> 
  summarise(nb_without_ambi = sum(is.na(GROUP)),
            nb_tot = n()) |> 
  mutate(percent_without_ambi_grp_taxon = nb_without_ambi / nb_tot * 100)

# Get list of species with no AMBI ecological group
grp_taxon_ambi |> 
  filter(is.na(GROUP)) |> 
  distinct(phylum, class, order, species) |> 
  distinct(species) |> arrange(species)
```

```{r}
ambi_index <- data_REBENT |> 
  mutate(species = case_when(
    species == "Nemertes" ~ "Nemertea",
    TRUE ~ species)) |> 
  group_by(ESTUARY, YEAR, tidal, species) |> 
  summarise(count = sum(RESULTAT)) |> 
  summarise(ambi = ambi(taxon = species, count = count), .groups = "drop")
```

```{r}
ggplot_ambi_index <- ggplot(ambi_index) +
  aes(x = YEAR, y = ambi, color = ESTUARY,
      linetype = tidal, shape = tidal) +
  geom_point() +
  geom_line() +
  labs(title = "REBENT macrobenthic fauna - AMBI", 
       x = NULL, y = "AZTI Marine Biotic Index",
       color = "Estuary")
ggplot_ambi_index
ggsave(plot = ggplot_ambi_index, filename = "../inst/results/data_benthos/indicators/ggplot_REBENT_AMBI_index.jpg",
       width = 15, height = 6, units = "cm")
```


## BEQI-FR

```{r}
ref_values <- tribble(
  ~ ESTUARY, ~ tidal, ~ EUNIS, ~ AMBI_ref, ~ H_ref, ~ S_ref,
  "Gironde", "intertidal", "A2.31", 2.5, 2.9, 14,
  "Gironde", "subtidal", "A5.32", 1.9, 2.5, 10,
  "Loire", "intertidal", "A2.31", 2.5, 2.9, 14,
  "Loire", "subtidal", "A5.25", 1.0, 3.8, 33,
  "Seine", "intertidal", "A2.24", 1.4, 3.7, 26,
  "Seine", "subtidal", "A5.22", 0.3, 2.7, 9
)
```

```{r}
BEQI <- species_richness |> 
  full_join(shannon_index) |> 
  full_join(ambi_index) |> 
  left_join(ref_values, by = c("ESTUARY", "tidal")) |> 
  mutate(EQR_ambi = (ambi - 7)/(AMBI_ref - 7),
         EQR_H = shannon / H_ref,
         EQR_S = sp_rich / S_ref,
         BEQIFR = (EQR_ambi + EQR_H + EQR_S)/3) |> 
  mutate(BEQIFR = case_when(
    BEQIFR >= 1 ~ 1,
    TRUE ~ BEQIFR
  ))
BEQI |> View()
```

```{r}
ggplot_BEQI_index <- ggplot(BEQI) +
  aes(x = YEAR, y = BEQIFR, color = ESTUARY,
      linetype = tidal, shape = tidal) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0.86) +
  geom_hline(yintercept = 0.67) +
  geom_hline(yintercept = 0.4) +
  labs(title = "REBENT macrobenthic fauna - BEQI-FR", 
       x = NULL, y = "Benthic Ecosystem Quality Index (BEQI-FR)",
       color = "Estuary") +
  annotate("text", x = 2015, y = 0.92, label = "High", size = 4, color = "black") +
  annotate("text", x = 2015, y = 0.76, label = "Good", size = 4, color = "black") +
  annotate("text", x = 2015, y = 0.53, label = "Moderate", size = 4, color = "black")

ggplot_BEQI_index
ggsave(plot = ggplot_BEQI_index, filename = "../inst/results/data_benthos/indicators/ggplot_REBENT_BEQIFR_index.jpg",
       width = 15, height = 6, units = "cm")
```
